---
title: "Canonical_all_niche_forms"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
execute: 
  cache: true
---

**Objectives:**
Check if the canonical form of the quadratic equation permit to fit different form of light spceies niche (concave, convex, positive or negative sigmoid, flat), when the model force a concave form of probabilities predictions.    

Canonical form: *a* \* (Environment - *O*)\^2 + *gamma*  

*a* = *beta2*\
*O* = extremum : -*beta1*/(2\**beta2)\
gamma = alpha*-(*beta1^2^/*(4*\**beta2))


```{r, packages, include = F}
library(rstan)
rstan_options(auto_write = TRUE) # option pour ne pas recompiler à chaque fois - gain de temps
options(mc.cores = parallel::detectCores()) #option pour ajouter des coeurs au calcul
rstan_options(disable_march_warning = TRUE)
library(bayesplot)
library(rstanarm) # for Bayesian automatic regression modelling using stan
library(brms) # Bayesian generalized multivariate non-linear multilevel models using stan
library(foreach) # parallisation
library(parallel)
library(tidyverse)
```

# Prepare data

```{r}
path <- "D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/"
```
sigmoid, convex or flat 
```{r, initial parameters}
dataParam <- data.frame(
  type = c("noeffect", "concave", "convexbell", "negsigmoid", "possigmoid"),
  alpha=0) %>% 
  mutate(beta1 = case_when(
    type == "noeffect" ~ 0,
    type == "concave" ~ 15,
    type == "convexbell" ~ -15,
    type == "negsigmoid" ~ -15,
    type == "possigmoid" ~ 15) 
  ) %>% 
  mutate(beta2 = case_when(
    type == "noeffect" ~ 0,
    type == "concave" ~ -15,
    type == "convexbell" ~ 15,
    type == "negsigmoid" ~ 0,
    type == "possigmoid" ~ 0) 
  ) %>% 
  mutate(a = beta2) %>% # Extremum
  mutate(O = -beta1/(2*beta2)) %>% # Extremum
  mutate(gamma = alpha-(beta1^2/(4*beta2)))

dataParamL <- dataParam %>% pivot_longer(cols = c("alpha", "beta1", "beta2", "a", "O", "gamma"),
                                         names_to = "Parameters",
                                         values_to = "Initial")

```

```{r, simulate data}
n <- 500 # abs + pres
X <- seq(0, 1, length.out = n) # env var
# Parameters
a <- 0 # intercept ; logit(0.5)= 0) autant de 0 que de 1

data <- list(
  # data.frame of simulated data
  noeffect = data.frame(b= 0,
                        c= 0, 
                        Environment = X) %>% 
    # Generate presence-absence data from probability
    mutate(Presence = rbinom(n, size = 1, # Bernoulli = 1 trial
                             # logistique transformation (inverse logit)
                             prob = 1 / (1 + exp(-(a + b * X + c * X^2))))),
  
  concave = data.frame(b= 15,
                       c= -15, 
                       Environment = X) %>% 
    # Generate presence-absence data from probability
    mutate(Presence = rbinom(n, size = 1, # Bernoulli = 1 trial
                             # logistique transformation (inverse logit)
                             prob = 1 / (1 + exp(-(a + b * X + c * X^2))))), 
  
  
  convexbell = data.frame(b=-15,
                          c= 15,
                          Environment = X) %>% 
    # Generate presence-absence data from probability
    mutate(Presence = rbinom(n, size = 1, # Bernoulli = 1 trial
                             # logistique transformation (inverse logit)
                             prob = 1 / (1 + exp(-(a + b * X + c * X^2))))), 
  
  negsigmoid = data.frame(b= -15,
                          c= 0,
                          Environment = X) %>% 
    # Generate presence-absence data from probability
    mutate(Presence = rbinom(n, size = 1, # Bernoulli = 1 trial
                             # logistique transformation (inverse logit)
                             prob = 1 / (1 + exp(-(a + b * X + c * X^2))))), 
  
  possigmoid = data.frame(b= 15,
                          c= 0,
                          Environment = X) %>% 
    # Generate presence-absence data from probability
    mutate(Presence = rbinom(n, size = 1, # Bernoulli = 1 trial
                             # logistique transformation (inverse logit)
                             prob = 1 / (1 + exp(-(a + b * X + c * X^2))))) 
)

# Visualisation of simulated data
for(x in names(data)){
  b <- unique(data[[x]]$b)
  c <- unique(data[[x]]$c)
  
  p <- ggplot(data[[x]], aes(x= Environment, y= Presence)) +
    geom_point() +
    theme_minimal() +
    stat_function(fun=function(x) 1 / (1 + exp(-(a + b * x + c * x^2))), col = "blue", linewidth = 1) +
    labs(title= x, y="Presence (1) / Absence (0)", x="Explicative variable (X)")
  
  print(p)
}

```

# Model
```{r, Model}
# quadratic - canonic form
M_Quadra_can <- stan_model(model_name = "Quadra_can",
                           paste(path, "Stanfiles/Bernoulli_EnvOnly_quadratic_canonic.stan", sep=''))

```

```{r, dataM}
Np <- 200 # number of predictions
Environmentp <- seq(0, 1, length.out = Np) # environment of predictions

dataM <- lapply(data, function(x) list(N = nrow(x), #les noms des var doivent etre les memes que dans le fichier stan
                                       Presence = x$Presence,
                                       Environment = x$Environment,
                                       # number of predictions
                                       Np = Np,
                                       # environment of predictions
                                       Environmentp = Environmentp)
)

names(dataM) <- names(data)
```

```{r, fit}
fits <- lapply(dataM, function(x) sampling(M_Quadra_can, data = x, chains = 4, iter = 2000))
names(fits) <- names(dataM)
```
-> There were **899 divergent transitions** after warmup.   
-> The largest **R-hat is 1.63, indicating chains have not mixed**.  
Running the chains for more iterations may help. Bulk Effective Samples Size (ESS) is too low, indicating **posterior means and medians may be unreliable**.  
Tail Effective Samples Size (ESS) is too low, indicating **posterior variances and tail quantiles may be unreliable**.  

-> There were 59 divergent transitions after warmup.   
Avis : The largest R-hat is 1.69, indicating chains have not mixed.  
Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.  
Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.  
-> There were 40 divergent transitions after warmup.  
-> The largest R-hat is 1.78, indicating chains have not mixed.  
Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.  
Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.  

# Chains

```{r, chains}
lapply(names(fits), function(x) mcmc_trace(as.array(fits[[x]]), 
                                           pars = c("a", "O", "gamma"),
                                           np = nuts_params(fits[[x]])) + # np pour afficher la divergeance
         labs(title = x)
)
```
-> apart from concave and convex, not all chains are mixed  

# Compare initial vs estimated parameters values

```{r, initial vs estimated parameters}
summaryfits <- lapply(names(fits), function(x) as.data.frame(summary(fits[[x]], 
                                                                     pars = c("a", "O", "gamma"))$summary[, c("50%", "2.5%", "97.5%")]) %>% 
                        tibble::rownames_to_column("Parameters")
)
names(summaryfits) <- names(fits)
summaryfits <- bind_rows(summaryfits, .id = "type")

summaryM <- dataParamL %>% 
  filter(Parameters %in% c("a", "O", "gamma")) %>% 
  left_join(summaryfits, by = c("type", "Parameters")) %>% 
  mutate(Difference = round(`50%`-Initial,2))

summaryM
```
Parameters estimation:    
-> noeffect: good  
-> concave: good  
-> convex: *a* is underestimated (-5,36), the rest is good.  
-> negsigmoid: *a* median is overestimated (6.28)  


# Parameters link

```{r, Parameters link}
lapply(names(fits), function(x) mcmc_pairs(as.array(fits[[x]]), pars = c("a", "O", "gamma")))
```
-> very bad except for concave an convex forms  

# Predictions
```{r, Predictions}
dataplot <- lapply(as.list(names(fits)), function(type)
  data.frame(type = type, 
             Environmentp = Environmentp,
             mu = apply(as.matrix(fits[[type]], pars = "p"), 2, median),
             t(apply(as.matrix(fits[[type]], pars = "p"), 2, 
                     quantile, probs = c(0.05, 0.95))),
        Opt = apply(as.matrix(fits[[type]], pars = "O"), 2, median),
        Opt = t(apply(as.matrix(fits[[type]], pars = "O"), 2,
                      quantile, probs = c(0.05, 0.95)))
             
  )) %>% 
  bind_rows()

ggplot(dataplot, aes(x = Environmentp)) +
  # Probas
  geom_point(aes(y = mu), size=0.7) +
  geom_ribbon(aes(ymin = `X5.`, ymax = `X95.`), color = 'red', alpha = 0.2) +
  labs(title = "Simulations median", y = "Presence probability", col= "Presence-absence") +
  facet_wrap(~ type, scales = "fixed", nrow = 1)

ggplot(dataplot, aes(x = Environmentp)) +
  # Probas
  geom_point(aes(y = mu), size=0.7) +
  geom_ribbon(aes(ymin = `X5.`, ymax = `X95.`), color = 'red', alpha = 0.2) +
  
  # Optimum
  geom_vline(aes(xintercept = Opt), color = "#00AFBB", linewidth=1) + # exp to delog
  geom_vline(aes(xintercept = `Opt.5.`), linetype="dashed") +
  geom_vline(aes(xintercept = `Opt.95.`), linetype="dashed") +
  labs(title = "Simulations median", y = "Presence probability", col= "Presence-absence") +
  facet_wrap(~ type, scales = "free_x", nrow = 1)

ggsave("Canonical_all_niche_forms.png", path = "D:/Mes Donnees/PhD/Figures/Modelisation", width = 35, height = 15, units = "cm", dpi=800, bg="white")
```

-> The distributions forms are well predicted   
-> The median extremum is interpretable is well positioned  
