---
title: "Convex - models selection"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
---

**Objectives:**
Essayer sur de vraies données convexes, si un modèle forcée concave est drastiquement moins bon qu’un modèle libre ou un modèle affine

Convex species: Licania_alba, Licania_membranacea, Bocoa_prouacensis, Iryanthera_sagotiana, Symphonia_sp.1, Pogonophora_schomburgkiana  


```{r, packages, include = F}
library(rstan)
rstan_options(auto_write = TRUE) # option pour ne pas recompiler à chaque fois - gain de temps
options(mc.cores = parallel::detectCores()) #option pour ajouter des coeurs au calcul
rstan_options(disable_march_warning = TRUE)
library(bayesplot)
library(rstanarm) # for Bayesian automatic regression modelling using stan
library(brms) # Bayesian generalized multivariate non-linear multilevel models using stan
library(foreach) # parallisation
library(parallel)
library(tidyverse)
library(loo)
```

# Prepare data

```{r, include = F, message=F, eval=F}
load("D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Realsp_4ha.Rdata")

sp <- c("Licania_alba", "Licania_membranacea", "Bocoa_prouacensis", "Iryanthera_sagotiana", "Symphonia_sp.1", "Pogonophora_schomburgkiana")

datalist <- datalist[names(datalist) %in% sp] # only species in sp
# view(datalist[["Licania_alba"]])
# data <- datalist[["Licania_alba"]]
# rm(datalist)
```

# Models
modèle forcée concave
modèle libre
modèle affine
```{r, eval=F}
path <- "D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/"

# quadratic - canonic form
M_Quadra_free <- stan_model(model_name = "Quadra_can_free",
                            paste(path, "Stanfiles/Bernoulli_EnvOnly_quadratic_canonic.stan", sep=''))

# quadratic - canonic form - concave forced
M_Quadra_forced <- stan_model(model_name = "Quadra_can_forced",
                              paste(path, "Stanfiles/Bernoulli_EnvOnly_quadratic_canonic_forced.stan", sep=''))

# modèle affine
M_linear <- stan_model(model_name = "linear",
                       paste(path, "Stanfiles/Bernoulli_EnvOnly_linear.stan", sep=''))

Models <- list(Affine= M_linear,
               Quadra_free = M_Quadra_free,
               Quadra_forced= M_Quadra_forced) # a list of models to parallelise

# rm(M_linear, M_Quadra_forced, Quadra_free)
```

# Run models
```{r, data list, eval=F}
# 6 species
Np <- 200 # number of predictions

dataM <- lapply(datalist, function(x) list(N = nrow(x), 
                                           Presence = x$Presence,
                                           Environment = x$logTransmittance,
                                           # number of predictions
                                           Np = Np,
                                           # environment of predictions
                                           Environmentp = seq(min(x$logTransmittance), max(x$logTransmittance), length.out = Np) 
)
)
# dataM
```

6 species * 3 models
```{r, eval=F}
# 18 models to run (6 sp * 3 model forms)
gc()
Sys.time() # 40 min
cores = min(length(Models), 5) # nbr of cores to use
i <- NULL
j = length(Models) # 3

for(S in sp) {
  # S=sp[1]
  
  # L'enregistrement des clusters
  cl <- parallel::makeCluster(cores, outfile = "")
  doSNOW::registerDoSNOW(cl)
  
  # Progress bar:
  pb <- txtProgressBar(max=j)
  progress <- function(n) cat(sprintf("Run %d is complete\n", n))
  opts <- list(progress = progress)
  
  fits <- foreach::foreach(
    i=1:j, # j
    .packages = c("rstan"), # necessary packages
    .options.snow = opts, # ProgressBar
    .combine=c) %dopar% {
      # the function to parallelise:
      sampling(Models[[i]], data = dataM[[S]], chains = 4, iter = 2000)
    }
  # close progressbar and cluster
  close(pb)
  stopCluster(cl)
  
  names(fits) <- names(Models[1:j])
  
  saveRDS(fits, file = paste("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Fits/ConvexSelecModel/QuadrafreevsforcedvsAffine",S,".rds", sep=""))
  
}

gc()
Sys.time()
```

-> Free :
There were 155 divergent transitions after warmup
The largest R-hat is 1.44, indicating chains have not mixed.
Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.

```{r, eval=F}
files <- list.files("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Fits/ConvexSelecModel",
                    pattern = "QuadrafreevsforcedvsAffine",full.names = T)#  files
# files
# 4 species and 4 models each

fits_list <- lapply(files, readRDS) # read files
names(fits_list) <- sp
names(fits_list[["Licania_alba"]])
```

```{r, include=F}
rm(M_linear,M_Quadra_forced, M_Quadra_free, fits, opts, pb, cores, files, i, j,S, progress); gc()
```

## Chains

```{r, eval=F}
mcmc_trace(as.array(Quadra_free), 
           pars = c("a", "O", "gamma"),
           np = nuts_params(Quadra_free)) +
  ggplot2::labs(title = "Quadratic free form")

mcmc_trace(as.array(Quadra_forced), 
           pars = c("a", "O", "gamma"),
           np = nuts_params(Quadra_forced)) +
  labs(title = "Quadratic concave forced form")

mcmc_trace(as.array(linear), 
           pars = c("alpha", "beta1"),
           np = nuts_params(linear)) +
  labs(title = "Linear form")
```

--> quadratic free form : divergences (pourquoi ?)

## Parameters link

```{r, eval=F}
mcmc_pairs(as.array(Quadra_free), pars = c("a", "O", "gamma"))

mcmc_pairs(as.array(Quadra_forced), pars = c("a", "O", "gamma"))

mcmc_pairs(as.array(linear), pars = c("alpha", "beta1"))
```

# Predictions
```{r, eval=F}
# form= "Quadra_dvp"
dataplot <- list()
for(S in names(fits_list)){
  dataplot[[S]] <- lapply(as.list(names(fits_list[[S]])), function(form)
    data.frame(form = form, 
               Environmentp = dataM[[S]]$Environmentp,
               mu = apply(as.matrix(fits_list[[S]][[form]], pars = "p"), 2, median),
               t(apply(as.matrix(fits_list[[S]][[form]], pars = "p"), 2, 
                       quantile, probs = c(0.05, 0.95))) #,
               # Opt = apply(as.matrix(fits_list[[form]], pars = "O"), 2, median),
               # Opt = t(apply(as.matrix(fits_list[[form]], pars = "O"), 2,
               #               quantile, probs = c(0.05, 0.95)))
               
    )) %>% 
    bind_rows()
}
View(dataplot[[S]])

plist <- vector('list')

for(S in names(fits_list)){
  
  plist[[S]] <- local({
    
    # print(
    #   ggplot(dataplot[[S]], aes(x = Environmentp)) +
    #     # Probas
    #     geom_point(aes(y = mu), size=0.7) +
    #     geom_ribbon(aes(ymin = `X5.`, ymax = `X95.`), color = 'red', alpha = 0.2) +
    #     labs(title = paste("Simulations median ",S," - y free"), y = "Presence probability", col= "Presence-absence") +
    #     facet_wrap(~ form, scales = "free", nrow = 1)
    # )
    
    print(
      ggplot(dataplot[[S]], aes(x = Environmentp)) +
        # Probas
        geom_point(aes(y = mu), size=0.7) +
        geom_ribbon(aes(ymin = `X5.`, ymax = `X95.`), color = 'red', alpha = 0.2) +
        labs(title =  paste("Simulations median ",S," - y fixed"), y = "Presence probability", col= "Presence-absence") +
        facet_wrap(~ form, scales = "fixed", nrow = 1)
    )
    
  })
}
plist

saveRDS(plist, file = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/Plots_convex.rds")
```

```{r}
readRDS("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/Plots_convex.rds")
```

# Model comparaison
```{r, loo, eval=F}
# ?loo
# class(fits_list[[1]][[1]])
# names(fits_list[[1]])
# Extract log likelihood

capture.output(
  
  for(S in names(fits_list)){
    # S="Licania_alba"
    
    print(S)
    
    log_lik_1 <- extract_log_lik(fits_list[[S]][["Quadra_free"]])
    log_lik_2 <- extract_log_lik(fits_list[[S]][["Quadra_forced"]])
    log_lik_3 <- extract_log_lik(fits_list[[S]][["Affine"]])
    
    # Compute LOO (leave-one-out cross-validation)
    # all Pareto k estimates have to be small (k< 0.5)
    Quadrafree <- loo(log_lik_1, is_method = "psis", cores=5) # Pareto-Smoothed Importance Sampling (PSIS)
    print(Quadrafree)
    # plot(Quadrafree)
    
    Quadraforced <- loo(log_lik_2, is_method = "psis", cores=5)
    print(Quadraforced)
    # plot(Quadraforced)
    
    affine <- loo(log_lik_3, is_method = "psis", cores=5)
    print(affine)
    # plot(affine)
    
    loos <- list(Quadrafree, Quadraforced, affine)
    names(loos) <- names(fits_list[[S]])
    
    # help('pareto-k-diagnostic') # for details.
    #elpd_loo = the expected log pointwise predictive density
    #p_loo = the effective number of parameters
    #looic = the LOO information criterion 
    
    
    # Model comparaison
    # ?loo_compare
    loo_diff <- loo_compare(loos)
    
    
    print(loo_diff, simplify=F)
    
  },
  
  file = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Loo/Loo_convex.txt")

```

```{r}
readLines("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Loo/Loo_convex.txt")
```

We can estimate the difference in their expected predictive accuracy by the difference in elpd_loo.  

The first row = model with the largest ELPD.  

A positive difference score like in this example, means that the second model is favored (the model with both predictors), whereas a negative score (elpd_diff) would indicate a preference for the first model.  

The standard error (se_diff) helps us decide if the difference in meaningful.  
If absolute value of the difference (elpd_diff) is less than the standard error (se_diff), the models don't perform differently, and we should choose the simpler model (the one with fewer parameters) because it is more parsimonious. If the absolute value of the difference is greater than the standard error, then we will prefer the model indicated by the test.  

"Licania_alba":  
-> first run: affine model is the best model. Quadradic forced is the worse model for convex. Now, Quadradic forced is the better model, and quadra free the worse.  

"Licania_membranacea":  
-> all models arge good, Affine is the best  

"Bocoa_prouacensis":  
-> affine model is the best model and the quadra free the worse.  

"Iryanthera_sagotiana":  
-> affine model is the best model  

"Symphonia_sp.1":  
-> all models arge good, Affine is the best 

"Pogonophora_schomburgkiana":  
-> affine model is the best model and the quadra free the worse with some K estimates too high.  

-> linear model is better for convex form. 
se_diff is close to elpd_diff absolute value so quadra forced is also a good model.  

```{r}

```

