---
title: "TreeHeight_measured_vs_estimated"
author: "Vincyane"
date: "`r Sys.Date()`"
output: html_document
---



# Giacomo FTH data
```{r, message=F}
FTH <- read_delim("~/PhD/Inventories/Data/Height-Diameter/Understory/Allometry_FTH_GSellan_Paracou_idTree.csv") %>% # 931
  select(ScientificName, Xutm, Yutm, DBH, TreeHeight) %>% 
  filter(!is.na(Xutm) & !is.na(Yutm) & !is.na(DBH)) %>% 
  rename(TreeHeight_m = TreeHeight)
  # mutate(dbh.y = as.numeric(gsub(",", ".", DBH)))
```

# Claudia's allometry - Michaelis-Menten LCH and species identity (MM4)
**//amap-data.cirad.fr/work/users/VincyaneBadouard/modèlesAlloClaudia**
```{r}
load("//amap-data.cirad.fr/work/users/VincyaneBadouard/modèlesAlloClaudia/model4_MM_dataCCmean.Rdata")
fit <- fit2_MM4_5s_2023
rm(fit2_MM4_5s_2023)

# model4_MM_nsing_CCmean <- readRDS("D:/Mes Donnees/PhD/Inventories/Data/Height-Diameter/CanopyConstructor/Claudia/model4_MM_nsing_CCmean.rds")
# View(model4_MM_nsing_CCmean$data)
```

# Scpecies correspondance
```{r}
match <- read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/Species_correspondence_Claudia.csv") # 399
```

```{r}
FTH <- FTH %>% 
  left_join(match, by= "ScientificName") %>% # add species_noSing
  rename(dbh.y = DBH)
```

# Predictions
## Hauteur médiane de canopée avoisinante (LCH)
```{r}
# P16 25ha CHM 1m res (SpatRaster)
CHM <- rast("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Paracou_P16_2023_25ha_50mbuffer_HighAlt_CHM_1mres.tif")

#  focal ("moving window") weight matrix   
disk_weights <- terra::focalMat(terra::rast(matrix(0, nrow=61, ncol=61)), 30, type = "circle") # 30m radius of the circle
# with target tree
disk_weights[disk_weights[]>0] <- 1

# Calculate focal ("moving window") values for each cell
CHMPP_median30cir <- focal(CHM, w=disk_weights, fun="median", na.policy="all",
                           fillvalue=NA, expand=FALSE, silent=TRUE)
# values in the focal window will be multiplied by the corresponding weight prior to 'fun' being applied.

plot(CHM) ; plot(CHMPP_median30cir)
```

```{r}
FTH_sf <- FTH %>% # as sf
  filter(!is.na(Xutm) & !is.na(Yutm)) %>% 
  st_as_sf(coords = c("Xutm", "Yutm")) %>% 
  st_set_crs(st_crs(CHM))

LCH <- raster::extract(CHMPP_median30cir, FTH_sf) %>% 
  rename(HC = focal_median) %>% dplyr::select(-ID)

FTH <- FTH %>% # add LCH to inventory
  cbind(LCH)

```

# Predict tree height for Claudia's species
```{r same vs other sp}
same_sp <- FTH %>% # 36673
  filter(!is.na(species_noSing)) %>% 
  mutate(type = "Specific")
  

other_sp <- FTH %>% 
  filter(is.na(species_noSing)) %>% # 11177
  mutate(type = "Community") 

nrow(FTH) == nrow(same_sp) + nrow(other_sp)
```

```{r prediction with posterior median}
same_sp_pred <- predict(fit, newdata = same_sp, robust=T, na.action = na.omit) # 1 min, robust=T to take the posterior median, mean if F
same_sp$TreeHeight_e <- same_sp_pred[,1]
```

# Predict tree height for species absent of the Claudia's selection (works)
si tu prends pas en compte l'effet de l'espèce, tu ne garde que l'effet moyen entre espèce (le mu de ta distribution des effets aléatoires des espèces). Donc tu considères toutes les espèces comme des espèces "moyennes"
```{r}
other_sp_pred <- predict(fit, newdata = other_sp, re_formula = NA, robust=T, na.action = na.omit) # 1 min; , robust=T to take the posterior median
other_sp$TreeHeight_e <- other_sp_pred[,1]
```

# Rebind & rename columns
```{r}
FTH_test <- same_sp %>%
  bind_rows(other_sp) %>% 
  rename(DBHcor = dbh.y) %>% 
  mutate(diff = TreeHeight_e - TreeHeight_m)


nrow(FTH_test) == nrow(FTH)
length(unique(FTH_test$ScientificName)) == length(unique(FTH$ScientificName)) 
```

# Plot
```{r}
# Compute median by group
median <- FTH_test %>% group_by(type)%>%summarise(Diff=median(diff, na.rm =T)) %>% na.omit()

FTH_test %>% filter(!is.na(diff)) %>% 
ggplot(aes(x= DBHcor, y = diff, group = type, col = type)) + # , group = type
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(data=median, aes(yintercept = Diff, col = type)) +
  geom_point(alpha=0.2) +
  geom_smooth() +
  labs(y = "Tree height estimated - measured (m)", x="DBH (cm)", col = 'Allometry used')
```

