---
title: "Light uncertainty tests"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
  markdown: 
    wrap: 72
code-fold: true
execute: 
  cache: true
---
Selected species with contrasted temperament and sufficiently abundant, with expert opinion:
- Tachigali melinonii (186, **helio**, canopy species)
- Anaxagorea dolichocarpa (938, **scia**, understory species)

```{r setup, include=F}
library(readr)
library(terra)
library(sf)
library(tidyverse)
library(tidybayes)
library(brms)
library(data.table)

library(lidR)
library(cmdstanr)
library(posterior)
library(factoextra)
library(grid)

library(bayesplot)
library(patchwork)
library(stringr)

sp <- c("Tachigali_melinonii","Anaxagorea_dolichocarpa")
N= 10
```

# Tree height uncertainty
**//amap-data.cirad.fr/work/users/VincyaneBadouard/modèlesAlloClaudia**
```{r Claudia fit}
load("//amap-data.cirad.fr/work/users/VincyaneBadouard/modèlesAlloClaudia/model4_MM_dataCCmean.Rdata")
fit <- fit2_MM4_5s_2023
rm(fit2_MM4_5s_2023)
# Formula: height ~ (alpha * dbh.y)/(beta + dbh.y) 
#          alpha ~ HC + 1 | species_noSing
#          beta ~ 1 | species_noSing
```

```{r plot posteriors, eval = F}
# "r_species_noSing__alpha[Tachigali.melinonii,Intercept]"  
# "r_species_noSing__beta[Tachigali.melinonii,Intercept]"

# plot_data <- fit %>%
#   spread_draws(r_species_noSing__alpha[Tachigali.melinonii,Intercept], r_species_noSing__beta[Tachigali.melinonii,Intercept]) %>%
#   filter(Tachigali.melinonii == "Tachigali.melinonii")
#   mutate(height = r_species_noSing__alpha /r_species_noSing__beta) # 
# 
# ggplot(plot_data, aes(y = `Tachigali.melinonii`, x = height)) +
#   ggdist::stat_halfeye()

plot(fit)
plot(fit, variable = c("r_species_noSing__alpha[Tachigali.melinonii,Intercept]", "r_species_noSing__beta[Tachigali.melinonii,Intercept]"))
```

```{r Import data, message=F}
Understory <- read.csv("~/PhD/Inventories/Data/Understory/Paracou/ALT_Paracou_9ha_CLEAN_corcord.csv") %>% 
  rename(idTree = IdTree) %>% # 36 311
  rename(SubPlot = subplot)
Adults <- read.csv("~/PhD/Inventories/Data/Adults/Paracou/AdultsP16_filtred.csv") %>%  # 25 ha 11 539
  mutate(idTree = as.character(idTree))

Inventory <- bind_rows(Understory, Adults)
```

```{r Modify inventory for predictions}
match <- read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/Species_correspondence_Claudia.csv")  # 399
# Generated with Modify_data_for_predictions.Rmd

Inventory <- Inventory %>% 
  left_join(match, by= "ScientificName") %>% # add species_noSing
  rename(dbh.y = DBHcor)
```

```{r Local canopy height (LCH)}
# P16 25ha CHM 1m res (SpatRaster)
CHM <- rast("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Paracou_P16_2023_25ha_50mbuffer_HighAlt_CHM_1mres.tif")

#  focal ("moving window") weight matrix   
disk_weights <- terra::focalMat(terra::rast(matrix(0, nrow=61, ncol=61)), 30, type = "circle") # 30m radius of the circle
# with target tree
disk_weights[disk_weights[]>0] <- 1

# Calculate focal ("moving window") values for each cell
CHMPP_median30cir <- focal(CHM, w=disk_weights, fun="median", na.policy="all",
                           fillvalue=NA, expand=FALSE, silent=TRUE)
# values in the focal window will be multiplied by the corresponding weight prior to 'fun' being applied.

Inventory_sf <- Inventory %>% # as sf
  filter(!is.na(Xutm) & !is.na(Yutm)) %>% 
  st_as_sf(coords = c("Xutm", "Yutm")) %>% 
  st_set_crs(st_crs(CHM))

LCH <- raster::extract(CHMPP_median30cir, Inventory_sf) %>% 
  rename(HC = focal_median) %>% dplyr::select(-ID)

Inventory <- Inventory %>% # add LCH to inventory
  cbind(LCH)
```

```{r same vs other sp}
same_sp <- Inventory %>% 
  filter(!is.na(species_noSing)) %>% # 36673
  mutate(Method_TreeHeight = "Species")

other_sp <- Inventory %>% 
  filter(is.na(species_noSing)) %>% # 11177
  mutate(Method_TreeHeight = "Community")
```

```{r H prediction with 100 posterior values - same sp}
same_sp_pred <- posterior_predict(fit, newdata = same_sp, ndraws = 10) # Génère des valeurs simulées de la variable réponse, 1 prédiction par draw du posterior
# rows : 1:10 predictions
# cols : 1:36673 trees

same_sp_pred <- same_sp_pred %>% 
  t() # inverse cols and rows

colnames(same_sp_pred)[1:10] <- paste0("TreeHeight_", 1:10)

same_sp <- bind_cols(same_sp, same_sp_pred)
```

```{r  H prediction with 100 posterior values - other sp}
other_sp_pred <- posterior_predict(fit, newdata = other_sp, re_formula = NA, ndraws = 10)

other_sp_pred <- other_sp_pred %>% 
  t() # inverse cols and rows

colnames(other_sp_pred)[1:10] <- paste0("TreeHeight_", 1:10)

other_sp <- bind_cols(other_sp, other_sp_pred)
```

```{r Bind preds}
pred <- same_sp %>%
  bind_rows(other_sp) %>% 
  rename(DBHcor = dbh.y)

nrow(pred) == nrow(Inventory)
length(unique(pred$ScientificName)) == length(unique(Inventory$ScientificName)) 
```

```{r Write data}
write.csv(pred, "D:/Mes Donnees/PhD/Inventories/Data/Adults_Understory/Paracou/AllP16_TreeHeights_incertitude.csv")
# pred <- read_csv("D:/Mes Donnees/PhD/Inventories/Data/Adults_Understory/Paracou/AllP16_TreeHeights_incertitude.csv") # 47,850

```

# Compute light for each tree height
```{r Read light file}
source("~/PhD/R_codes/PhD/LiDAR_scripts/Functions/ReadLightFile.R")
path <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2019/Light/P16_2019_25ha_buffer_Light_grid_intensity2m.txt"

Light <- ReadLightFile(path, skiplines=9)

setnames(Light, "Period 1", "Transmittance")

Light <- Light[Transmittance>0 & Z>2,] # remove Z=2m because all is equal to 100%T (trees zutm >= 5.7)
```

```{r Extract Zutm for each tree}
data <- pred %>% 
  select(idTree, ScientificName, Strategy, Xutm, Yutm, SubPlot, DBHcor, contains("TreeHeight"))

MNT <- terra::rast("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/MNT/dtm2023_25ha_HighAlt_buffer.asc") 
terra::crs(MNT) <- terra::crs("EPSG:2972")

datasf <- st_as_sf(data, coords = c('Xutm','Yutm'))
datasf <- st_set_crs(datasf, st_crs(MNT)) 

alt <- terra::extract(MNT, datasf) %>% # extract z
  rename(Zutm = `MNT_0.5m_IRD_500m_Paracou_284000_579500`)

data <- data %>% cbind(alt) %>% setDT()
# data
```

```{r Coordinates in local positions (use VOP) as light file}
VOP <- as.matrix(read.table("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/VOP/VOP_P16_25ha.txt"))

data <- setDT(data)

matrix <- as.matrix(data[, `:=`(c = 0 , d = 1)][,.(Xutm, Yutm, c, d)]) # as matrix of the same dimensions, without Z

datalocal <- matrix %*% t(VOP) # same dimensions

datalocal <- as.data.frame(cbind(datalocal, data[,Zutm]))[, -c(3,4)]  # add Zutm
names(datalocal) <- c("X","Y","Z")
datalocal <- datalocal %>%
  # cbind(data[,.-()]) %>%
  cbind(data[,c("SubPlot", "idTree", "ScientificName", "Strategy", "Xutm", "Yutm", "Zutm", "DBHcor",
                names(data)[grepl("TreeHeight", names(data))] ), with = FALSE])
# datalocal
```

```{r Pivot}
datalocal <- datalocal %>% 
  pivot_longer(cols = starts_with("TreeHeight"),
               names_to = "Sample",
               values_to = "TreeHeight") %>%
  setDT()
```

Modification : obliger la hauteur de l'arbre a un minimum d'1m30 (hauteur de mesure du diamètre)
```{r negative heights}
neg_H <- datalocal %>% 
  filter(TreeHeight<0)
range(neg_H$DBHcor) # 1 3.5
length(unique(neg_H$idTree)) # 6151
length(unique(neg_H$idTree))/length(unique(datalocal$idTree))*100
dupli <- neg_H %>% 
  group_by(idTree) %>% mutate(N = n()) # max 7/10 hauteurs estimées négatives

# TreeHeight is 1.3 minimum
datalocal <- datalocal %>%
  mutate(TreeHeight = ifelse(TreeHeight<1.3, 1.3, TreeHeight))
```

```{r Coordinates 1m above each tree}
datalocal[,Z := Z+TreeHeight+1] # zutm + tree height +1
range(datalocal$Z) # 3.906876 68.748097

datalocal[Z>max(Light$Z),] #> 66m, 11 tree height estimations too tall

# datalocal <- datalocal[Z<max(Light$Z),] # remove trees too tall


# datalocal
```

```{r}
summary(Light[, .(X, Y, Z)]) # max z=66
summary(datalocal[, .(X, Y, Z)]) # max tree height too big 68.9m
```
**changement : j'ai augmenté le radius de 4 à 6 pour avoir 8 voisins pour toutes les hauteurs d'arbres**
```{r Light per tree height}
# Get neighbors in immediate neighborhood (k  nearest voxels including target)
result <- RANN::nn2(data = Light[, .(X, Y, Z)], # Pool to search
                    query = datalocal[, .(X, Y, Z)], # [1,], # point of interest
                    # The maximum number of nearest neighbours to compute
                    k = 8, # including target # 5, 8, 9, 15, 27, 30, 35, 35
                    # !!!mettre 8 (pas plus pour ne pas "tartiner"
                    searchtype = "radius", radius = 6) # (4) pour éviter les NA en bord de scene 
# k = 30, radius = 3
# k = 7, radius = 5

# une matrice où chaque ligne correspond à un point et contient les indices de ses voisins les plus proches
index <- result$nn.idx # the near neighbour indices
any(index==0)

# Light <- tibble::rownames_to_column(Light, "Rowname")
# Light[Rowname %in% c(8586286, 9077687, 8586285, 8585585, 8586987, 8586287, 9077686, 9076986, 9078388, 8094885, 9077688, 8585584, 8586986, 8585586, 8586988, 9076985, 9078387, 8094884, 8094184, 9076987, 8095586, 9078389, 8094886, 9569088, 8094183, 8095585, 8094185, 8095587, 8586284, 9569087),]

# Compute light from local neighborhood :
# each row of index is a point of interest

fill.value <- apply(index, 1, function(i) Light[as.vector(i), mean(Transmittance)] # for a matrix 1 indicates rows
)
any(fill.value==0 | is.na(fill.value))

datalocal$Transmittance <- fill.value  
```

Inv (idTree, ScientificName, Xutm, Yutm, Zutm, DBHcor, TreeHeight) + Transmittance
```{r Write data}
datalocal <- datalocal %>%
  select(idTree, SubPlot, ScientificName, Strategy, Xutm, Yutm, Zutm, DBHcor, Sample, TreeHeight, Method_TreeHeight, Z, Transmittance)

write.csv(datalocal, "~/PhD/Inventories/Data/Adults_Understory/Paracou/AllP16_Light_2019_1m_incertitude.csv") 
# datalocal <- read.csv("~/PhD/Inventories/Data/Adults_Understory/Paracou/AllP16_Light_2019_1m_incertitude.csv")
```

# Add TWI
```{r}
Inv_env <- read_csv("D:/Mes Donnees/PhD/Inventories/Data/Adults_Understory/Paracou/P16_2019_Paracou_InvandEnv.csv") %>% 
  filter(SubPlot %in% c(13,14,15,18,19,20,23,24,25)) %>% 
  select(idTree,SubPlot,ScientificName,TWI)


datalocal <- datalocal %>% 
  filter(SubPlot %in% c(13,14,15,18,19,20,23,24,25)) %>% 
  left_join(Inv_env, by= c('idTree','SubPlot','ScientificName'))

any(is.na(datalocal$TWI))

datalocal <- datalocal %>% 
  mutate(logTransmittance =  log(Transmittance)) %>% 
  mutate(logTWI = log(TWI))

write.csv(datalocal, "D:/Mes Donnees/PhD/Inventories/Data/Adults_Understory/Paracou/P16_2019_Paracou_InvandEnv_incertitude.csv")
```

# Presence-absence data
```{r Presence-absence}
Inv <- datalocal

# Create a list of the inventory replicated, 1 per ScientificName
datalist <- replicate(length(sp), Inv, simplify = F) 
names(datalist) <- sp # 1 per ScientificName

# Presence=1 if it's the interest species, 0 if not
datalist <- lapply(names(datalist), function(name) {
  datalist[[name]]$Presence <- ifelse(
    !is.na(datalist[[name]]$ScientificName) & datalist[[name]]$ScientificName == name, 1, 0)
  
  datalist[[name]]  # Return modified data frame
})
names(datalist) <- sp 
```

```{r}
save(datalist, file= "D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Realsp_9ha_incertitude.Rdata")

# load("D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Realsp_9ha_incertitude.Rdata")
```

# Join eigenvectors
```{r}
filelist <- list.files("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/Per_species_9ha/", full.names=T)
eigenval <- lapply(filelist, read_csv)
filelist <- gsub("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/Per_species_9ha/", "", filelist)
filelist <- gsub(".csv", "", filelist)

names(eigenval) <- filelist
eigenval <- eigenval[names(eigenval) %in% sp] # eigen val per tree position
sp <- names(eigenval)
```

```{r}
eigenval <- lapply(names(datalist), function(name) {
  eigenval[[name]] <- eigenval[[name]] %>%
    slice(rep(row_number(), each = 10))
})
names(eigenval) <- sp

datalist_eig <- datalist
# str(datalist_eig[[1]])

datalist_eig <- lapply(names(datalist), function(name) {
  # name = "Sextonia_rubra"
  if(name %in% names(eigenval) & ncol(eigenval[[name]])>0){
    datalist_eig[[name]] <- datalist_eig[[name]] %>%
      # left_join(unique(eigenval[[name]]), by=c("Xutm", "Yutm"), copy = T)
      bind_cols(eigenval[[name]])
    
  }
  datalist_eig[[name]]  # Return modified data frame
})

names(datalist_eig) <- names(datalist) 
```

```{r}
save(datalist_eig, file= "D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Realsp_9ha_eigenvectors_incertitude.Rdata")

# load("D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Realsp_9ha_eigenvectors_incertitude.Rdata")
```

# Fits
```{r fit}
Comb <- read.csv("D:/Mes Donnees/PhD/R_codes/PhD_cluster/Data/Combin_Sp_Sample.csv")

PATH <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/Light_uncertainty"

fits <- list()
for(S in sp){
  # # Filter species with only one DBH class
  # if(length(list.files(file.path("Chains", "Hybrid_eigenvectors", S)))>1){
  
  for(D in paste0("TreeHeight_", 1:N)){
    tryCatch({
      chain_path <- file.path("Chains", "Hybrid_eigenvectors", S, D) 
      
      fits[[S]][[D]] <- as_cmdstan_fit(list.files(chain_path,
                                                  full.names = TRUE))},
      error=function(e){cat("ERROR :",S, D, conditionMessage(e), "\n")}
    )
  }
  # }
}
# names(fits[["Tachigali_melinonii"]])


sp[!sp %in% names(fits)] # 

saveRDS(fits, file = paste(PATH, "/fits.rds", sep=''))
# fits <- readRDS(paste(PATH, "/fits.rds", sep=''))
```

# Display uncertainty
combiner la dernière valeur du posterior de chaque chaîne (car autocorrélation intrachaine, convergence sur une valeur) des SDM de chaque estimation de hauteur d’une espèce pour avoir la variance de toute l’incertitude pour l’espèce (100 x 4 chaines = 400 valeurs → distribution de l’incertitude)
```{r}
d <- list()
for(S in names(fits)){
  for(D in paste0("TreeHeight_", 1:N)){
    if(D %in% names(fits[[S]])){
      
      d[[S]][[D]] <- fits[[S]][[D]]$draws(c("a", "O"), format = "df") %>% 
        filter(.iteration == 2000) %>% 
        mutate(Species = S) %>% 
        mutate(Sample = D) %>% 
        select(-.draw)
      
    }}}
draws <- list_rbind(list_flatten(d))

draws <- draws %>%
  mutate(Species = recode(Species,
                          "Anaxagorea_dolichocarpa" = "Tachigali_melinonii" ,
                          "Tachigali_melinonii" = "Anaxagorea_dolichocarpa"))
```

```{r}
initialfits <- readRDS("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/Light_topo/Autocor/fits.rds")

initialfits <- initialfits[names(initialfits) %in% sp]

d <- list()
for(S in names(initialfits)){
  
  d[[S]] <- initialfits[[S]]$draws(c("a", "O"), format = "df") %>% 
    filter(.iteration == 2000) %>% 
    mutate(Species = S) %>% 
    select(-.draw)
  
}
initialdraws <- list_rbind(list_flatten(d))
```

```{r}
ggplot(draws, aes(x= O)) +
  theme_minimal() +
  geom_histogram(binwidth=1, aes(fill="Uncertainty propagation"), color="#e9ecef", alpha=0.9) +
  geom_histogram(data = initialdraws, aes(fill="Posterior median"),
                 binwidth=1, color="#e9ecef", alpha=0.9) +
  scale_x_continuous(n.breaks = 5, labels = ~round(exp(.)*100,2)) + # delog
  facet_wrap(~ Species) +
  scale_fill_manual(values = c("Posterior median" = "#009E73", # pour donner un nom à chaque intitulé
                               "Uncertainty propagation" = "#D55E00"),
                    breaks=c("Posterior median", "Uncertainty propagation")) + # pour ordonner
  theme(legend.position="bottom") +
  labs(title="Tree height uncertainty propagation", fill = "Tree height")
```
```{r}
range(initialdraws[initialdraws$Species == "Anaxagorea_dolichocarpa",]$O) # -4.18563 -4.04452
range(draws[draws$Species == "Anaxagorea_dolichocarpa",]$O) # -4.28972 -3.81780

range(initialdraws[initialdraws$Species == "Tachigali_melinonii",]$O) # -1.400090 -0.643009
range(draws[draws$Species == "Tachigali_melinonii",]$O) # -2.5436500 -0.0224763

draws %>% 
  group_by(Species) %>% 
  summarise(Uncertainty = max(O) - min(O))

initialdraws %>% 
  group_by(Species) %>% 
  summarise(Uncertainty_initial = max(O) - min(O))

# delog
draws %>% 
  group_by(Species) %>% 
  summarise(Uncertainty = (exp(max(O))*100) - (exp(min(O))*100))

initialdraws %>% 
  group_by(Species) %>% 
  summarise(Uncertainty_initial = (exp(max(O))*100) - (exp(min(O))*100))


# exp(-4.04452)
# exp(-1.400090) # 0.2465748
# exp(-0.643009) # 0.5257082
# 
# exp(-2.5436500) # 0.07857906
# exp(-0.0224763) # 0.9777744
```


```{r}
ggplot(draws, aes(x= a)) +
  theme_minimal() +
  geom_histogram(binwidth=.2, aes(fill="Uncertainty propagation"), color="#e9ecef", alpha=0.9) +
  geom_histogram(data = initialdraws, aes(fill="Posterior median"),
                 binwidth=.2, color="#e9ecef", alpha=0.9) +
  facet_wrap(~ Species) +
  scale_fill_manual(values = c("Posterior median" = "#009E73", # pour donner un nom à chaque intitulé
                               "Uncertainty propagation" = "#D55E00"),
                    breaks=c("Posterior median", "Uncertainty propagation")) + # pour ordonner
  theme(legend.position="bottom") +
  labs(title="Tree height uncertainty propagation", fill = "Tree height")
```
## Posteriors with O deloged and -a logged
```{r}
datap_initial <- read_csv("//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/Light_topo/Autocor/dataplot.csv") %>% 
  filter(ScientificName %in% sp)
```

```{r}
# trace(bayesplot::mcmc_intervals_data, edit=T)
probs <- c(0.5 - 0.8/2, 0.5 - 0.5/2, 0.5 + 0.5/2, 0.5 + 0.8/2)

datap <- draws %>% 
  group_by(Species) %>% 
  summarise(across(c("O", "a"),
                   list(m = ~ median(.),
                        ll =  ~ quantile(., probs[1]),
                        l =  ~ quantile(., probs[2]),
                        h =  ~ quantile(., probs[3]),
                        hh = ~ quantile(., probs[4]))))

a <- datap %>% 
  select(-starts_with("O_"))
a <- a %>% 
  rename_with(.fn = ~gsub("a_", "", names(a)), .cols = names(.)) %>% 
  mutate(parameter = 'a') 

O <- datap %>% 
  select(-starts_with("a_"))
O <- O %>% 
  rename_with(.fn = ~gsub("O_", "", names(O)), .cols = names(.)) %>% 
  mutate(parameter = 'O')
datap <- O %>% 
  bind_rows(a) %>% 
  mutate(Source = 'Propagation')%>% 
  mutate(Species = paste0(gsub("_", " ", Species), "\npropagation"))
```

Ajouter les posterieurs de la propagation d'incertitude 
```{r}
# 0 line
x_lim <- range(c(datap_initial$ll, datap_initial$hh))
x_range <- diff(x_lim)
x_lim[1] <- x_lim[1] - 0.05 * x_range
x_lim[2] <- x_lim[2] + 0.05 * x_range

# 0 line
layer_vertical_line <- if (0 > x_lim[1] && 0 < x_lim[2]) {
  vline_0(color = "darkred", linewidth = 0.5, linetype = "dashed")
}

datap_ord <- datap_initial %>% 
  mutate(parameter = factor(parameter,
                            levels=c("O","a", "tau")
  )) %>% 
  group_by(parameter) %>%
  group_modify(~.x %>% 
                 # arrange(if(.y == 'tau') m else desc(m))
                 arrange(if(.y == 'O') m else desc(m))
  ) %>%
  ungroup() %>%
  mutate(Species=factor(Species, levels=unique(Species))) %>% 
  mutate(Species = paste0(Species, "\ninitial")) %>% 
  bind_rows(datap)

xs <- split(datap_ord, f = datap_ord$parameter)

# Residuals_covar %>% 
p1 <- xs$O %>% 
  
  # mutate(Source = 'initial') %>% 
  # bind_rows(O) %>% 
  ggplot(aes(y = Species)) + # , alpha = Source, group = Source
  theme_minimal() + 
  # xlim(range(c(datap$ll, datap$hh))) +
  layer_vertical_line +
  geom_segment(aes(x=ll, xend=hh, y=Species, yend=Species,
                   color = "80%"), size = 0.5) +
  geom_segment(aes(x=l, xend=h, y=Species, yend=Species,
                   color = "50%"), size = 1) +
  geom_point(aes(x= m)) + # median
  scale_color_manual(name = "Credible interval", 
                     values = c("80%" = "#E7B800",
                                "50%"= "#FC4E07")) + 
  scale_alpha_manual(name = "Source", values=c(1,.2),
                     breaks=c("initial","Propagation"))+
  theme(axis.title.x =element_blank())

p2 <- p1 %+% xs$a +
  ggtitle("a") +
  # scale_x_continuous(trans="log") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p1 <- p1 + ggtitle("O (% transmittance)") +
  scale_x_continuous(n.breaks = 5, labels = ~round(exp(.)*100,2)) + # delog
  guides(color = FALSE, alpha = FALSE)


p1+p2

# ggsave("Light_uncertainty_posteriors_delog.pdf", path = PATH,
#        width = 25, height = 26, units = "cm", dpi=800, bg="white")

```

# Eigenvectors per sample
```{r}
filelist <- list.files("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/TreeHeight_uncertainty/Tachigali_melinonii/", full.names=T)
eigenval_s <- lapply(filelist, read_csv)
filelist <- gsub("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/TreeHeight_uncertainty/Tachigali_melinonii/", "", filelist)
filelist <- gsub(".csv", "", filelist)

names(eigenval_s) <- filelist

# how many eigenvectorrs in common
sum(names(eigenval[["Tachigali_melinonii"]]) %in% names(eigenval_s[[10]])) # 2-5/5 in common

names(eigenval[["Tachigali_melinonii"]])
names(eigenval_s[[1]])

# ------------------------------------------------------------------------------
filelist <- list.files("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/TreeHeight_uncertainty/", pattern="*.csv", full.names=T)
eigenval_s <- lapply(filelist, read_csv)
filelist <- gsub("D:/Mes Donnees/PhD/Inventories/Data/Agregation/Eigenvectors/TreeHeight_uncertainty/", "", filelist)
filelist <- gsub(".csv", "", filelist)

names(eigenval_s) <- filelist

# how many eigenvectorrs in common
sum(names(eigenval[["Anaxagorea_dolichocarpa"]]) %in% names(eigenval_s[[10]])) # 5/5 in common

names(eigenval[["Anaxagorea_dolichocarpa"]])
names(eigenval_s[[1]])
```

