---
title: "Diagnose all"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
---

```{r setup}
library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(patchwork)

sp <- (read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/InterestSpecies.csv")[,3]) # 75
# mettre 2 pour treedatacor

# sp <- c("Licania_membranacea" , "Lecythis_persistens") # 4 forms
# , "Symphonia_sp.1", "Tachigali_melinonii",
#         "Iryanthera_hostmannii", "Tabernaemontana_macrocalyx", "Pouteria_singularis", "Ryania_speciosa"
# sp <- "Symphonia_sp.1"

strategy <- read_delim("~/PhD/Inventories/Data/Understory/Paracou/Species_strategy.csv") %>% 
  select(ScientificName, Strategy_VB) %>%
  rename(Strategy = Strategy_VB) %>% 
  unique()
```

```{r fit}
fits <- list()
for(S in sp){
  tryCatch({
    chain_path <- file.path("Chains", "Hybrid", S)
    fits[[S]] <- as_cmdstan_fit(list.files(chain_path,
                                           full.names = TRUE))},
    error=function(e){cat("ERROR :",S, conditionMessage(e), "\n")}
  )
}
# names(fits[["Symphonia_sp.1"]])
```

```{r divergence}
Div <- lapply(names(fits), function(x)
  data.frame(Species = x,
             Chain = c(1:4),
             Divergences = fits[[x]]$diagnostic_summary("divergences") # divergences per chain
  ) 
) %>% 
  bind_rows() %>% 
  rename(Divergences = num_divergent) %>% 
  group_by(Species) %>% 
  mutate(`%divergence` = sum(Divergences)/4000*100)

```

# Rhat
R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1.01 (Vehtari et al., 2021). 
R-hat which is the maximum of rank normalized split-R-hat and rank normalized folded-split-R-hat, which works for thick tailed distributions and is sensitive also to differences in scale.
```{r Rhat}
Rhat_table <- lapply(names(fits), function(x)
  fits[[x]]$summary(c("alpha", "beta1", "beta2_p", "tau")) %>% 
    mutate(Species = x) %>% 
    select(Species, variable, rhat) 
) %>% 
  bind_rows()

# bayesplot::rhat(fits[[1]],pars="beta2_p")
# posterior::rhat(extract_variable_matrix(fits[[1]], "beta2_p"))

all(Rhat_table$rhat < 1.01)

badRhat <- Rhat_table %>% filter(rhat > 1.01) %>%
  left_join(Div %>% select(Species, `%divergence`) %>% unique(), by="Species")
```

```{r}
name <- names(fits)#  unique(badRhat$Species) # names(fits) #
lapply(name , function(x)
  fits[[x]]$draws(c("alpha", "beta1", "beta2_p", "tau")) %>% 
    mcmc_trace() +
    ggtitle(x)
)
```

```{r}
# no light effect (flat) :
no_light_effect <- c("Iryanthera_sagotiana", "Symphonia_sp.1", "Inga_loubryana", "Virola_michelii", "Pouteria_guianensis", "Eschweilera_congestiflora", "Couratari_multiflora", "Tachigali_melinonii", "Pouteria_gonggrijpii", "Thyrsodium_puberulum", "Anacardium_spruceanum")

# a_Q5 < -0.07
# (a = 0 -> large niche)
```

# Posteriors
Plot central (quantile-based) posterior interval estimates from MCMC draws
```{r}
# Plot data
datap <- lapply(names(fits), function(x)
  mcmc_intervals_data(fits[[x]]$draws(c("a", "O", "tau")),
                      prob = 0.5, prob_outer = 0.95, point_est = "median") %>% 
    mutate(Species = x)
)  %>% 
  bind_rows() %>% 
  # tau significantly different of 0
  mutate(Sg_effect = ifelse(parameter=="tau" & ll<0 & 0<hh, F, T))  %>% 
  # Light effect (not flat)
  mutate(Sg_effect = ifelse(parameter=="O" & Species %in% no_light_effect, F, Sg_effect)) 

# ll : outer start
# l : iner start
# m : point_est value
# h : iner end
# hh : outer end

# mcmc_intervals() code 
# trace(bayesplot::mcmc_intervals, edit=T)

# 0 line
x_lim <- range(c(datap$ll, datap$hh))
x_range <- diff(x_lim)
x_lim[1] <- x_lim[1] - 0.05 * x_range
x_lim[2] <- x_lim[2] + 0.05 * x_range

# 0 line
layer_vertical_line <- if (0 > x_lim[1] && 0 < x_lim[2]) {
  vline_0(color = "darkred", linewidth = 0.5, linetype = "dashed")
}
datap %>% 
  mutate(parameter = factor(parameter,
                            levels=c("tau", "O", "a")
                            # levels=c("O","a","tau")
  )) %>% 
  group_by(parameter) %>%
  group_modify(~.x %>% 
                 arrange(if(.y == 'tau') m else desc(m))
               # arrange(if(.y == 'O') m else desc(m))
  ) %>%
  ungroup() %>%
  mutate(Species=factor(Species, levels=unique(Species))) %>% 
  ggplot(aes(y = Species, alpha = Sg_effect)) +
  theme_minimal() + 
  # xlim(range(c(datap$ll, datap$hh))) +
  layer_vertical_line +
  geom_segment(aes(x=ll, xend=hh, y=Species, yend=Species,
                   color = "95%"), size = 0.5) +
  geom_segment(aes(x=l, xend=h, y=Species, yend=Species,
                   color = "50%"), size = 1) +
  geom_point(aes(x= m)) + # median
  scale_color_manual(name = "Credible interval", 
                     values = c("95%" = "#E7B800",
                                "50%"= "#FC4E07")) + 
  scale_alpha_manual(name = "Non-null effect", values=c(1,.2),
                     breaks=c("TRUE","FALSE")) +
  theme(axis.title.x =element_blank()) +
  facet_grid(~parameter, scale= "free")

gridExtra::grid.arrange()

ggsave("Light_topo_posteriors.pdf", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 25, height = 26, units = "cm", dpi=800, bg="white")

# O et tau iront dans le biplot
# les paramètres dont on veut la significativité : iota
# a on veut sa valeur absolue pour juger la largeur de niche
# gamma on s'en fout ?
```

## Posteriors with O deloged
```{r}
# Plot data
datap <- lapply(names(fits), function(x)
  mcmc_intervals_data(fits[[x]]$draws(c("a", "O", "tau")),
                      prob = 0.5, prob_outer = 0.95, point_est = "median") %>% 
    mutate(Species = x)
)  %>% 
  bind_rows() %>% 
  # tau significantly different of 0
  mutate(Sg_effect = ifelse(parameter=="tau" & ll<0 & 0<hh, F, T))  %>% 
  # Light effect (not flat)
  mutate(Sg_effect = ifelse(parameter=="O" & Species %in% no_light_effect, F, Sg_effect)) 

# ll : outer start
# l : iner start
# m : point_est value
# h : iner end
# hh : outer end

# mcmc_intervals() code 
# trace(bayesplot::mcmc_intervals, edit=T)

# 0 line
x_lim <- range(c(datap$ll, datap$hh))
x_range <- diff(x_lim)
x_lim[1] <- x_lim[1] - 0.05 * x_range
x_lim[2] <- x_lim[2] + 0.05 * x_range

# 0 line
layer_vertical_line <- if (0 > x_lim[1] && 0 < x_lim[2]) {
  vline_0(color = "darkred", linewidth = 0.5, linetype = "dashed")
}
datap_ord <- datap %>% 
  mutate(parameter = factor(parameter,
                            # levels=c("tau", "O", "a")
                            levels=c("O","a","tau")
  )) %>% 
  group_by(parameter) %>%
  group_modify(~.x %>% 
                 # arrange(if(.y == 'tau') m else desc(m))
                 arrange(if(.y == 'O') m else desc(m))
  ) %>%
  ungroup() %>%
  mutate(Species=factor(Species, levels=unique(Species))) 

xs <- split(datap_ord, f = datap_ord$parameter)

# Residuals_covar %>% 
p1 <- xs$O %>% 
  ggplot(aes(y = Species, alpha = Sg_effect)) +
  theme_minimal() + 
  # xlim(range(c(datap$ll, datap$hh))) +
  layer_vertical_line +
  geom_segment(aes(x=ll, xend=hh, y=Species, yend=Species,
                   color = "95%"), size = 0.5) +
  geom_segment(aes(x=l, xend=h, y=Species, yend=Species,
                   color = "50%"), size = 1) +
  geom_point(aes(x= m)) + # median
  scale_color_manual(name = "Credible interval", 
                     values = c("95%" = "#E7B800",
                                "50%"= "#FC4E07")) + 
  scale_alpha_manual(name = "Non-null effect", values=c(1,.2),
                     breaks=c("TRUE","FALSE"))+
  theme(axis.title.x =element_blank())

p2 <- p1 %+% xs$a +
  ggtitle("a") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  guides(color = FALSE, alpha = FALSE)

p3 <- p1 %+% xs$tau  +
  ggtitle("tau") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p1 <- p1 + ggtitle("O (% transmittance)") +
  scale_x_continuous(n.breaks = 5, labels = ~round(exp(.)*100,2)) + # delog
  guides(color = FALSE, alpha = FALSE)


p1+p2+p3

ggsave("Light_topo_posteriors_delog.pdf", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 25, height = 26, units = "cm", dpi=800, bg="white")

```

```{r}
fits[[1]]$draws(c("a", "O", "gamma", "tau")) %>% mcmc_areas_ridges()
```

# Biplot O vs tau
O in y, tau in x
```{r}
# color and alpha background for topo and light env
rects <- data.frame(xstart = sort(rep(c(-7.5, .5, -0.5), times=2)), xend = sort(rep(c(-0.5, .5, 3), times=2)),
                    ystart = rep(c(-7,log(0.5)), times =3), yend = rep(c(log(0.5),0), times =3), 
                    col = rep(c("Hilltop", "No effect", "Bottomland"), each =2),
                    light = rep(c("Sciaphile", "Heliophile"), times =3))
```

```{r}
databp <- datap %>% 
  select(-c(outer_width, inner_width, point_est)) %>% 
  rename(Q5 = ll, Q95 = hh, median = m, Q50_0 = l, Q50_1 = h) %>% 
  filter(parameter %in% c("O", "tau")) %>%
  select(-Sg_effect) %>% 
  pivot_wider(names_from = parameter, values_from = c(Q5, Q50_0, median, Q50_1, Q95),
              names_glue = "{parameter}_{.value}") %>% 
  # tau significantly different of 0
  mutate(Tau_effect = ifelse(tau_Q5<0 & 0<tau_Q95, F, T))  %>% 
  # Light effect (not flat)
  mutate(Sg_light_effect = ifelse(Species %in% no_light_effect, F, T)) %>% 
  left_join(strategy, by = c("Species" = "ScientificName"))

exp(quantile(databp$O_median, probs = 0.5))*100


(nrow(databp %>% filter(Tau_effect))/nrow(databp))*100 # 66% sp with tau significantly (95% CI) different of 0

# hc <- hclust(dist(databp[c("tau_median", "O_median")])) # Hierarchical Clustering
# plot(hc)
# databp$cluster <- cutree(hc, k = 4) # Cut a Tree into Groups
```

## Complete
```{r}
ggplot(databp) +
  theme_minimal() + 
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin=xstart,
                            xmax=xend, fill=col, alpha=light)) +
  scale_fill_manual(name = "Topographic preference",
                    values = c("Hilltop" = "#E7B800",
                               "Bottomland"= "steelblue",
                               "No effect"= "grey"),
                    breaks = c("Hilltop","No effect","Bottomland")) + 
  
  scale_alpha_manual(name = "Light preference", values=c(.2,.5)) +
  
  # geom_smooth(aes(x= tau_median, y = O_median), method='lm', formula= y~x, col = "#999999", alpha = 0.2) +
  ggpubr::stat_cor(aes(x= tau_median, y = O_median), label.x = -7.5, label.y = .1) +
  
  geom_vline(xintercept = 0, col="red", linetype = "dashed") +
  
  # 95% credible interval
  geom_segment(aes(x=tau_Q5, xend=tau_Q95, y=O_median, yend=O_median,
                   col = "95 and 50% CI"), size = .5) + # tau
  geom_segment(aes(y=O_Q5, yend=O_Q95, x=tau_median, xend=tau_median,
                   col = "95 and 50% CI"), size = .5) + # O
  geom_segment(aes(x=tau_Q50_0, xend=tau_Q50_1, y=O_median, yend=O_median,
                   col = "95 and 50% CI"), size = .6) + # tau
  geom_segment(aes(y=O_Q50_0, yend=O_Q50_1, x=tau_median, xend=tau_median,
                   col = "95 and 50% CI"), size = .6) + # O
  
  
  geom_point(aes(x= tau_median, y = O_median, # median
                 col=Tau_effect, shape = Sg_light_effect)) + 
  scale_shape_manual(name="Light effect",
                     values=c("FALSE" = 1, "TRUE" = 19),
                     breaks = c("TRUE","FALSE")) +
  
  scale_color_manual(name = "Significant topographic effect",
                     values = c("FALSE" = "grey",
                                "TRUE"= "black",
                                "95 and 50% CI" = "darkgrey"),
                     breaks=c("TRUE",  "FALSE", 
                              "95 and 50% CI")) + 
  
  ggrepel::geom_text_repel(aes(x= tau_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  # stat_ellipse(aes(x= tau_median, y = O_median, color = as.factor(cluster)),
  #              type = "t",geom = "polygon",alpha = 0.4) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  # scale_y_continuous(trans="exp") +
  guides(fill = guide_legend(nrow=2), alpha = guide_legend(nrow=2),
         col = guide_legend(nrow=2), shape = guide_legend(nrow=2)) +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") + # , legend.box="vertical"
  labs(x= "Topographic Wetness Index (TWI) linear effect (*tau*)",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("Light_topo_biplot_complete.png", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```

## No error bar
```{r}
ggplot(databp) +
  theme_minimal() + 
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin=xstart,
                            xmax=xend, fill=col, alpha=light)) +
  scale_fill_manual(name = "Topographic preference",
                    values = c("Hilltop" = "#E7B800",
                               "Bottomland"= "steelblue",
                               "No effect"= "grey"),
                    breaks = c("Hilltop","No effect","Bottomland")) + 
  
  scale_alpha_manual(name = "Light preference", values=c(.2,.5)) +
  
  # geom_smooth(aes(x= tau_median, y = O_median), method='lm', formula= y~x, col = "#999999", alpha = 0.2) +
  ggpubr::stat_cor(aes(x= tau_median, y = O_median), label.x = -7.5, label.y = .1) +
  
  geom_vline(xintercept = 0, col="red", linetype = "dashed") +
  
  geom_point(aes(x= tau_median, y = O_median, # median
                 col=Tau_effect, shape = Sg_light_effect)) + 
  scale_shape_manual(name="Light effect",
                     values=c("FALSE" = 1, "TRUE" = 19),
                     breaks = c("TRUE","FALSE")) +
  
  scale_color_manual(name = "Significant topographic effect",
                     values = c("FALSE" = "grey",
                                "TRUE"= "black",
                                "95 and 50% CI" = "darkgrey"),
                     breaks=c("TRUE",  "FALSE", 
                              "95 and 50% CI")) + 
  
  ggrepel::geom_text_repel(aes(x= tau_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  # stat_ellipse(aes(x= tau_median, y = O_median, color = as.factor(cluster)),
  #              type = "t",geom = "polygon",alpha = 0.4) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  # scale_y_continuous(trans="exp") +
  guides(fill = guide_legend(nrow=2), alpha = guide_legend(nrow=2),
         col = guide_legend(nrow=2), shape = guide_legend(nrow=2)) +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") + # , legend.box="vertical"
  labs(x= "Topographic Wetness Index (TWI) linear effect (*tau*)",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("Light_topo_biplot_noerrorbar.png", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```

## Without no light effect
```{r}
databp %>% 
  filter(Sg_light_effect) %>% 
  ggplot() +
  theme_minimal() + 
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin=xstart,
                            xmax=xend, fill=col, alpha=light)) +
  scale_fill_manual(name = "Topographic preference",
                    values = c("Hilltop" = "#E7B800",
                               "Bottomland"= "steelblue",
                               "No effect"= "grey"),
                    breaks = c("Hilltop","No effect","Bottomland")) + 
  
  scale_alpha_manual(name = "Light preference", values=c(.2,.5)) +
  
  # geom_smooth(aes(x= tau_median, y = O_median), method='lm', formula= y~x, col = "#999999", alpha = 0.2) +
  ggpubr::stat_cor(aes(x= tau_median, y = O_median), label.x = -7.5, label.y = .1) +
  
  geom_vline(xintercept = 0, col="red", linetype = "dashed") +
  
  geom_point(aes(x= tau_median, y = O_median, # median
                 col=Tau_effect)) + 
  
  scale_color_manual(name = "Significant topographic effect",
                     values = c("FALSE" = "grey",
                                "TRUE"= "black",
                                "95 and 50% CI" = "darkgrey"),
                     breaks=c("TRUE",  "FALSE", 
                              "95 and 50% CI")) + 
  
  ggrepel::geom_text_repel(aes(x= tau_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  # stat_ellipse(aes(x= tau_median, y = O_median, color = as.factor(cluster)),
  #              type = "t",geom = "polygon",alpha = 0.4) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  # scale_y_continuous(trans="exp") +
  guides(fill = guide_legend(nrow=2), alpha = guide_legend(nrow=2),
         col = guide_legend(nrow=2), shape = guide_legend(nrow=2)) +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") + # , legend.box="vertical"
  labs(x= "Topographic Wetness Index (TWI) linear effect (*tau*)",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("Light_topo_biplot_noerrorbar_onlysg_light_effect.png", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```

# Biplot O vs a

O in y, tau in x
```{r}
# color and alpha background for light env
rects <- data.frame(xstart = sort(rep(c(-7.5, 0.5, 0), times=2)), xend = sort(rep(c(-0.5, 0, 3), times=2)),
                    ystart = rep(c(-7,log(0.5)), times =3), yend = rep(c(log(0.5),0), times =3), 
                    light = rep(c("Sciaphile", "Heliophile"), times =3))
```

```{r}
databp <- datap %>% 
  select(-c(outer_width, inner_width, point_est)) %>% 
  rename(Q5 = ll, Q95 = hh, median = m, Q50_0 = l, Q50_1 = h) %>% 
  filter(parameter %in% c("O", "a")) %>% 
  pivot_wider(names_from = parameter, values_from = c(Q5, Q50_0, median, Q50_1, Q95),
              names_glue = "{parameter}_{.value}") %>% 
  # Light effect (not flat)
  mutate(Sg_light_effect = ifelse(Species %in% no_light_effect, F, T)) 


# hc <- hclust(dist(databp[c("a_median", "O_median")])) # Hierarchical Clustering
# plot(hc)
# databp$cluster <- cutree(hc, k = 2) # Cut a Tree into Groups
```

```{r}
ggplot(databp) +
  theme_minimal() + 
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin= min(databp$a_Q5),
                            xmax=0, alpha=light), fill="grey") +
  # scale_fill_manual(name = "Topographic preference", values = c("Hilltop" = "#E7B800",
  #                                                               "Bottomland"= "steelblue",
  #                                                               "No effect"= "white")) + 
  scale_alpha_manual(name = "Light preference", values=c(.1,.5)) +
  
  # geom_smooth(aes(x= tau_median, y = O_median), method='lm', formula= y~x, col = "#999999", alpha = 0.2) +
  ggpubr::stat_cor(aes(x= a_median, y = O_median), label.x = -.7, label.y = .1) +
  
  # 95% credible interval
  geom_segment(aes(x=a_Q5, xend=a_Q95, y=O_median, yend=O_median,
                   col = "95 and 50% CI"), size = .5) + # tau
  geom_segment(aes(y=O_Q5, yend=O_Q95, x=a_median, xend=a_median,
                   col = "95 and 50% CI"), size = .5) + # O
  geom_segment(aes(x=a_Q50_0, xend=a_Q50_1, y=O_median, yend=O_median,
                   col = "95 and 50% CI"), size = .6) + # tau
  geom_segment(aes(y=O_Q50_0, yend=O_Q50_1, x=a_median, xend=a_median,
                   col = "95 and 50% CI"), size = .6) + # O
  
  
  geom_point(aes(x= a_median, y = O_median, shape = Sg_light_effect)) + # median
  scale_shape_manual(name="Light effect",
                     values=c("FALSE" = 1, "TRUE" = 19),
                     breaks = c("TRUE","FALSE")) +
  scale_color_manual(name="", values = c("95 and 50% CI" = "darkgrey")) + 
  
  ggrepel::geom_text_repel(aes(x= a_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  # stat_ellipse(aes(x= a_median, y = O_median, color = as.factor(cluster)),
  #              type = "t",geom = "polygon",alpha = 0.4) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  # scale_y_continuous(trans="exp") +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") +
  labs(x= "Niche width (*a*)",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("O_a_biplot.png", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```

# Predictions
```{r predictions}
Sys.time()

plist <- list()

for(s in names(fits)){
  plist[[s]] <- local({
    # s= "Lecythis_persistens"
    
    DATA <- fits[[s]]$summary("p") %>%  # p posterior
      # prediction environment 
      rowwise() %>% 
      mutate(Light = as.numeric(unlist(str_extract_all(variable, "\\d+"))[1])) %>% 
      mutate(Topo = as.numeric(unlist(str_extract_all(variable, "\\d+"))[2])) %>% 
      mutate(logTransmittance = seq(-7.308724e+00, 8.941571e-08, length.out = 50)[Light]) %>% # -7.308724e+00 , 8.941571e-08
      mutate(logTWI = seq(1.358045, 2.400176, length.out = 50)[Topo]) # 1.358045 - 2.400176
    
    DATA %>% 
      group_by(Light) %>% 
      # At topo optimum (i.e. presence proba. max)
      filter(median==max(median)) %>% 
      ggplot(aes(x = logTransmittance)) +
      theme_minimal() +
      geom_point(aes(y = median), size=0.7) + # p posterior median
      geom_ribbon(aes(ymin = q5, ymax = q95), color = 'red', alpha = 0.2) + # P quantiles
      geom_vline(xintercept  = fits[[s]]$summary("O", median)$median, color = "#00AFBB", linewidth=1) + # Opt median
      geom_vline(xintercept  = fits[[s]]$summary("O")$q5, linetype="dashed") + # Opt Q5%
      geom_vline(xintercept  = fits[[s]]$summary("O")$q95, linetype="dashed") + # Opt Q95%
      # scale_x_continuous(trans="log") +
      scale_x_continuous(name = "Light (% transmittance)",
                         n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
      ggtitle(paste(s,"- at logTWI optimum")) +
      ylab("Presence probabiblity median")
    
  })
}


plots <- gridExtra::marrangeGrob(plist, nrow = 3, ncol = 3) # 75
# plots

ggsave("Allsp_Light_niche_atTWIoptimum.pdf",
       path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       plots, width = 15, height = 10)

Sys.time() # 32 min
```

# Strategies boxplot
```{r}
ggplot(databp, aes(x= Strategy, y= O_median)) +
  geom_boxplot()
```


