---
title: "Diagnose all"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
---

```{r setup}
library(tidyverse)
library(cmdstanr)
library(bayesplot)

sp <- (read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/InterestSpecies.csv")[,3]) # 75
# mettre 2 pour treedatacor

# sp <- c("Licania_membranacea" , "Lecythis_persistens") # 4 forms
# , "Symphonia_sp.1", "Tachigali_melinonii",
#         "Iryanthera_hostmannii", "Tabernaemontana_macrocalyx", "Pouteria_singularis", "Ryania_speciosa"
# sp <- "Symphonia_sp.1"

```

```{r fit}
fits <- list()
for(S in sp){
  tryCatch({
    chain_path <- file.path("Chains", "Hybrid", S)
    fits[[S]] <- as_cmdstan_fit(list.files(chain_path,
                                           full.names = TRUE))},
    error=function(e){cat("ERROR :",S, conditionMessage(e), "\n")}
  )
}
# names(fits[["Symphonia_sp.1"]])
```

```{r divergence}
Div <- lapply(names(fits), function(x)
  data.frame(Species = x,
             Chain = c(1:4),
             Divergences = fits[[x]]$diagnostic_summary("divergences") # divergences per chain
  ) 
) %>% 
  bind_rows() %>% 
  rename(Divergences = num_divergent) %>% 
  group_by(Species) %>% 
  mutate(`%divergence` = sum(Divergences)/4000*100)

```

# Rhat
R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1.01 (Vehtari et al., 2021). 
R-hat which is the maximum of rank normalized split-R-hat and rank normalized folded-split-R-hat, which works for thick tailed distributions and is sensitive also to differences in scale.
```{r Rhat}
Rhat_table <- lapply(names(fits), function(x)
  fits[[x]]$summary(c("alpha", "beta1", "beta2_p", "tau")) %>% 
    mutate(Species = x) %>% 
    select(Species, variable, rhat) 
) %>% 
  bind_rows()

# bayesplot::rhat(fits[[1]],pars="beta2_p")
# posterior::rhat(extract_variable_matrix(fits[[1]], "beta2_p"))

all(Rhat_table$rhat < 1.01)

badRhat <- Rhat_table %>% filter(rhat > 1.01) %>%
  left_join(Div %>% select(Species, `%divergence`) %>% unique(), by="Species")
```

```{r}
name <-names(fits)#  unique(badRhat$Species) # names(fits) #
lapply(name , function(x)
  fits[[x]]$draws(c("alpha", "beta1", "beta2_p", "tau")) %>% 
    mcmc_trace() +
    ggtitle(x)
)
```

# Posteriors
Plot central (quantile-based) posterior interval estimates from MCMC draws
```{r}
# Plot data
datap <- lapply(names(fits), function(x)
  mcmc_intervals_data(fits[[x]]$draws(c("a", "O", "tau")),
                      prob = 0.5, prob_outer = 0.95, point_est = "median") %>% 
    mutate(Species = x)
)  %>% 
  bind_rows()
# ll : outer start
# l : iner start
# m : point_est value
# h : iner end
# hh : outer end

# mcmc_intervals() code 
# trace(bayesplot::mcmc_intervals, edit=T)

# 0 line
x_lim <- range(c(datap$ll, datap$hh))
x_range <- diff(x_lim)
x_lim[1] <- x_lim[1] - 0.05 * x_range
x_lim[2] <- x_lim[2] + 0.05 * x_range

# 0 line
layer_vertical_line <- if (0 > x_lim[1] && 0 < x_lim[2]) {
  vline_0(color = "darkred", linewidth = 0.5, linetype = "dashed")
}
datap %>% 
  mutate(parameter = factor(parameter, levels=c("O","a","tau"))) %>% # "tau", "O", "a"
  group_by(parameter) %>%
  group_modify(~.x %>% 
                 arrange(if(.y == 'O') m else desc(m)) ) %>%
  ungroup() %>%
  mutate(Species=factor(Species, levels=unique(Species))) %>% 
  ggplot(aes(y = Species)) +
  theme_minimal() + 
  # xlim(range(c(datap$ll, datap$hh))) +
  layer_vertical_line +
  geom_segment(aes(x=ll, xend=hh, y=Species, yend=Species,
                   color = "95%"), size = 0.5) +
  geom_segment(aes(x=l, xend=h, y=Species, yend=Species,
                   color = "50%"), size = 1) +
  geom_point(aes(x= m)) + # median
  scale_color_manual(name = "Credible interval", 
                     values = c("95%" = "#E7B800",
                                "50%"= "#FC4E07")) + 
  theme(axis.title.x =element_blank()) +
  facet_grid(~parameter, scale= "free")


ggsave("Light_topo_posteriors.pdf", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 25, height = 26, units = "cm", dpi=800, bg="white")

# O et tau iront dans le biplot
# les paramètres dont on veut la significativité : iota
# a on veut sa valeur absolue pour juger la largeur de niche
# gamma on s'en fout ?
```


```{r}
fits[[1]]$draws(c("a", "O", "gamma", "tau")) %>% mcmc_areas_ridges()
```

# Biplot
O in y, tau in x
```{r}
rects <- data.frame(xstart = sort(rep(c(-7.5, 0.5, 0), times=2)), xend = sort(rep(c(-0.5, 0, 3), times=2)),
                    ystart = rep(c(-7,log(0.5)), times =3), yend = rep(c(log(0.5),0), times =3), 
                    col = rep(c("Hilltop", "No effect", "Bottomland"), each =2),
                    light = rep(c("Sciaphile", "Heliophile"), times =3))
datap %>% 
  select(-c(outer_width, inner_width, l, h, point_est)) %>% 
  rename(Q5 = ll, Q95 = hh, median = m) %>% 
  filter(parameter %in% c("O", "tau")) %>% 
  pivot_wider(names_from = parameter, values_from = c(Q5, median, Q95),
              names_glue = "{parameter}_{.value}") %>% 
  mutate(Noeffect = ifelse(tau_Q5<0 & 0<tau_Q95, F, T)) %>% 
  
  ggplot() +
  theme_minimal() + 
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin=xstart,
                            xmax=xend, fill=col, alpha=light)) +
  scale_fill_manual(name = "Topographic preference", values = c("Hilltop" = "#E7B800",
                                                                "Bottomland"= "steelblue",
                                                                "No effect"= "white")) + 
  scale_alpha_manual(name = "Light preference", values=c(0.2,0.5)) +
  
  # geom_smooth(aes(x= tau_median, y = O_median), method='lm', formula= y~x, col = "#999999", alpha = 0.2) +
  ggpubr::stat_cor(aes(x= tau_median, y = O_median), label.x = -7.5, label.y = 0.1) +
  
  geom_vline(xintercept = 0, col="red", linetype = "dashed") +
  
  # 95% credible interval
  geom_segment(aes(x=tau_Q5, xend=tau_Q95, y=O_median, yend=O_median,
                   col = "95% CI"), size = 0.5) + # tau
  geom_segment(aes(y=O_Q5, yend=O_Q95, x=tau_median, xend=tau_median,
                   col = "95% CI"), size = 0.5) + # O
  
  geom_point(aes(x= tau_median, y = O_median, col=Noeffect)) + # median
  scale_color_manual(name = "*tau* significantly different of 0",
                     values = c("FALSE" = "grey",
                                "TRUE"= "black",
                                "95% CI" = "darkgrey"),
                     breaks=c("TRUE",  "FALSE", 
                              "95% CI")) + 
  
  ggrepel::geom_text_repel(aes(x= tau_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  # scale_y_continuous(trans="exp") +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") +
  labs(x= "Topographic Wetness Index (TWI) linear effect (*tau*)",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("Light_topo_biplot.png", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```


# Predictions
```{r predictions}
Sys.time()

plist <- list()

for(s in sp){
  plist[[s]] <- local({
    
    g <- lapply(names(fits[[s]]), function(v){
      # v = "TWI"
      
      DATA <- fits[[s]][[v]]$summary("p") # p posterior
      # prediction environment 
    })
    
  })
}

plots <- gridExtra::marrangeGrob(plist, nrow = 3, ncol = 3) # 75
# plots

# ggsave("Allsp_Light_TWI_niche.pdf",
#        path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
#        plots, width = 15, height = 10)

Sys.time()
```


