---
title: "Data_Tomst_visualisation_script"
author: "Vincyane Badouard"
date: "`r Sys.Date()`"
output: pdf_document

---

# Working directory
Change the path if your work with another folder
```{r, setup}
# Folder path
path0 <- "D:/Mes Donnees/PhD/Humidity - data and scripts/LollyData/Nouragues_Test_2023_02_22"

data_94244920_2023_02_22_0 <- read.csv("~/PhD/Humidity - data and scripts/LollyData/Nouragues_Test_2023_02_22/data_94244920_2023_02_22_0.csv", header=FALSE, sep=";")
```

# Data formatting function
Columns :
- index number of the measure,
- date and time in UTC,
- time zone = 4 (French Guiana is UTC-3)
- T1, T2, T3,
- soil moisture count,
- shake = 0
- errFlag (if =1 the device couldn’t convert time from PCF chip) = NA
```{r}
# data <- tms1n
format_tms <- function(
    # data, # data name (character) with ".csv"
    path = path0 # folder path
){
  
  # Data name
  data0 <- data 
  
  # Subset data
  data <- read.csv2(file.path(path, data), h = F); # row 1 is not the headers
  data <- data[3360:dim(data)[1],] # start at the 3360 rown (before it's jerome's office T°)
  
  # The date
  date1 <- strptime(data$V2,"%Y.%m.%d %H:%M")
  # Convert time in UTC-3 (ToDO)
  
  # Temperature data
  temperature1 <- data$V4
  temperature2 <- data$V5
  temperature3 <- data$V6
  
  # Moisture data
  moisture <- data$V7
  
  # Bind columns
  data2 <- cbind(as.data.frame(date1),
                 as.data.frame(temperature1), # soil T°
                 as.data.frame(temperature2), # surface T°
                 as.data.frame(temperature3), # air T°
                 as.data.frame(moisture))
  
  Output <- list("data" = data2,# dataset
                 "file" = data0, # data name
                 "length" = length(temperature1)) # length of eah dataset
  
  return(Output)
}
```


```{r}
format_tms <- function(
    data_path # path of the data (character) with ".csv"
){
  
  # Data name
  file_name <- data_path 
  
  # Subset data
  data <- read.csv2(data_path, h = F); # row 1 is not the headers
  # data <- data[3360:dim(data)[1],] # start at the 3360 rown (before it's jerome's office T°)
  
  # The date
  date1 <- strptime(data$V2,"%Y.%m.%d %H:%M")
  # Convert time in UTC-3 (ToDO)
  
  # Temperature data
  temperature1 <- data$V4
  temperature2 <- data$V5
  temperature3 <- data$V6
  
  # Moisture data
  moisture <- data$V7
  
  # Bind columns
  data2 <- cbind(as.data.frame(date1),
                 as.data.frame(temperature1), # soil T°
                 as.data.frame(temperature2), # surface T°
                 as.data.frame(temperature3), # air T°
                 as.data.frame(moisture))
  
  Output <- list("data" = data2,# dataset
                 "file_name" = file_name, 
                 "length" = length(temperature1)) # length of eah dataset
  
  return(Output)
}
```

# Temperature plot function
```{r}
# Plot de la temperature a trois points (air=10 > sol; sol=niveau du sol, deep= 10 cm < sol)
plottmsT <- function(
    tms,
    colabove, # air T° colour
    coltop, # surface T° colour
    coldeep # soil T° colour
){
  
  #rangex=(2000:2400)
  rangex <- (1:tms$length)
  
  # Air T° 
  plot(tms$data$date1[rangex],
       tms$data$temperature3[rangex],
       col = colabove,
       ylim = range(tms$data[,2:4]), pch = 20, cex = 0.3, type = "l",
       xlab = "Day", ylab = "Temperature (C)")
  
  title(main = paste("File name:",tms$file),
        sub = paste("Mean temp: Soil=", round(mean(tms$data$temperature1),2),
                    " Surface=", round(mean(tms$data$temperature2),2),
                    " Air=", round(mean(tms$data$temperature3),2)))
  
  # Ground T°
  lines(tms$data$date1,
        tms$data$temperature2,
        col = coltop, pch = 20)
  
  # soil T°
  lines(tms$data$date1,tms$data$temperature1,col = coldeep, pch = 20)
  
  #points(tms$date1,tms$temperature2, col = coldeep, pch = 20,cex = 0.5)
  #points(tms$date1, tms$temperature3, col = colabove, pch = 20, cex = 0.5)
  
}

```

# Humidity plot function
```{r}
# plot de l'humidite a 10 cm de profondeur
plottmsM <- function(
    tms,
    col # colour
){
  
  plot(tms$data$date1,
       tms$data$moisture,
       col = col, pch = 20,cex = 0.3, type = "b",
       xlab = "Day", ylab = "Moisture SU")
  
  title(main = paste("File name:", tms$file),
        sub = paste("Mean moisture=", round(mean(tms$data$moisture),2))
  )
  
}
```

# Data load and format
```{r}
#Lecture et formattage des fichiers (modifie au 16/12/2021, donnees jusqu'au 10/12/2021)
tms1n <- "data_94244901_2023_02_22_0.csv";tms1 = format_tms(tms1n, path0)
tms2n <- "data_94244902_2023_02_22_0.csv";tms2 = format_tms(tms2n)
tms3n <- "data_94244903_2023_02_22_0.csv";tms3 = format_tms(tms3n)
tms4n <- "data_94244904_2023_02_22_0.csv";tms4 = format_tms(tms4n)
tms5n <- "data_94244905_2023_02_22_0.csv";tms5 = format_tms(tms5n)
tms6n <- "data_94244906_2023_02_22_0.csv";tms6 = format_tms(tms6n)
tms7n <- "data_94244907_2023_02_22_0.csv";tms7 = format_tms(tms7n)
tms8n <- "data_94244908_2023_02_22_0.csv";tms8 = format_tms(tms8n)
tms9n <- "data_94244909_2023_02_22_0.csv";tms9 = format_tms(tms9n)
tms10n <- "data_94244910_2023_02_22_0.csv";tms10 = format_tms(tms10n)
tms11n <- "data_94244911_2023_02_22_0.csv";tms1 = format_tms(tms11n)
tms12n <- "data_94244912_2023_02_22_0.csv";tms2 = format_tms(tms12n)
tms13n <- "data_94244913_2023_02_22_0.csv";tms3 = format_tms(tms13n)
tms14n <- "data_94244914_2023_02_22_0.csv";tms4 = format_tms(tms14n)
tms15n <- "data_94244915_2023_02_22_0.csv";tms5 = format_tms(tms15n)
tms16n <- "data_94244916_2023_02_22_0.csv";tms6 = format_tms(tms16n)
tms17n <- "data_94244917_2023_02_22_0.csv";tms7 = format_tms(tms17n)
tms18n <- "data_94244918_2023_02_22_0.csv";tms8 = format_tms(tms18n)
tms19n <- "data_94244919_2023_02_22_0.csv";tms9 = format_tms(tms19n)
tms20n <- "data_94244920_2023_02_22_0.csv";tms10 = format_tms(tms20n)

```

```{r}
path <- "D:/Mes Donnees/PhD/Humidity - data and scripts/LollyData/Nouragues_Test_2023_02_22"
files <- list.files(path, pattern = "data", full.names = T)[1:20] # 20 the number of files
data <- lapply(files, format_tms)
names(data) <- files
bind_rows(data, .id = "file") %>% 
  unnest() %>% 
  mutate(file = gsub(paste0(path, "/"), "", file)) %>% 
  mutate(date = lubridate::as_datetime(date1)) %>% 
  select(date, temperature1, temperature2, temperature3, moisture) %>% 
  reshape2::melt("date") %>% 
  ggplot(aes(date, value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~variable, scales = "free_y") +
  theme_bw()
```

# Plot temperature
```{r}
par(mfrow=c(2,1))
# plot all temperatures (blue=air, orange=surface, brown=soil)
dev.copy(device = pdf,
         file="temporal_trends_temperature.pdf",
         width = 7, height = 7, pointsize = 12)

plottmsT(tms1,"blue","orange","brown")
plottmsT(tms2,"blue","orange","brown")
plottmsT(tms3,"blue","orange","brown")
plottmsT(tms4,"blue","orange","brown")
plottmsT(tms5,"blue","orange","brown")
plottmsT(tms6,"blue","orange","brown")
plottmsT(tms7,"blue","orange","brown")
plottmsT(tms8,"blue","orange","brown")
plottmsT(tms9,"blue","orange","brown")
plottmsT(tms10,"blue","orange","brown")

dev.off()

```

# Plot soil moisture
```{r}
dev.copy(device = pdf,
         file="temporal_trends_moisture.pdf", # output file name
         width = 7, height = 7, pointsize = 12)

plottmsM(tms1,"blue")
plottmsM(tms2,"blue")
plottmsM(tms3,"blue")
plottmsM(tms4,"blue")
plottmsM(tms5,"blue")
plottmsM(tms6,"blue")
plottmsM(tms7,"blue")
#plottmsM(tms8,"blue")
#plottmsM(tms9,"blue")
plottmsM(tms10,"blue")
dev.off()
```

