---
title: "Stakes_coordinates"
author: "Vincyane"
date: "`r Sys.Date()`"
output: html_document
---
# Packages
```{r Setup, include = F}
library(readr)
library(readxl)
library(tidyverse)
```

# Import data
Site : Paracou - P16
Subplots :
- 13, 18, 23, 24 : oui
- 14-15-19-20-25 : pas encore 
Méthode de mesure : laser

```{r, 13_18_23_24}
# 13, 18, 23, 24
S_13_18_23_24 <- read_delim("~/PhD/Inventories/Data/Piquets_ALT/Measures_piquets_Paracou13_18_23_24.csv", 
                              delim = ";", escape_double = FALSE, trim_ws = TRUE) # 401
View(S_13_18_23_24)
names(S_13_18_23_24) # "plot" "quadra.nb.X" "quadra.nb.Y" "q.length.X"  "q.length.X2" "q.length.Y"  "q.length.Y2"
summary(S_13_18_23_24) # Min q.length.X : 1.0
S_13_18_23_24 <- S_13_18_23_24[complete.cases(S_13_18_23_24),] # enlever les lignes à NA -> 400
```
"plot" : subplot number
"quadra.nb.X" : number of column where the 10m x 10m quadrat is placed (0 - 9. 0 = sud; 9 = north).  
"quadra.nb.Y" : number of row where the 10m x 10m quadrat is placed (0 - 9. 0 = west; 9 = east).  
"q.length.X" : size of the southern side of the 10m x 10m quadrat (m).  
"q.length.X2" : size of the northern side of the 10m x 10m quadrat (m).  
"q.length.Y" : size of the western side of the 10m x 10m quadrat (m).  
"q.length.Y2" : size of the eastern side of the 10m x 10m quadrat (m).  

```{r}
S_14 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C14_measure.xlsx")
S_15 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C15_measure.xlsx")
S_19 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C19_measure.xlsx")
S_20_1 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C20_measure.xlsx")
S_20_2 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C20_fin20250130.xlsx")
S_25 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/Measures_piquets_ParacouC25_T1_20230607ins.xlsx")


all(names(S_14) == names(S_15))
all(names(S_14) == names(S_19))
names(S_20_1) == names(S_14)
names(S_14) == names(S_20_2) == names(S_25)

names(S_14) # "Num" "Q_lengthX" "Q_lengthX1" "Q_lengthY" "Q_lengthY1" "Subplot" "Comments"
summary(S_14) # Min q.length.X : 1.0

```
# Manage column names differences
```{r}
S_13_18_23_24 <- S_13_18_23_24 %>% 
  rename(Subplot = plot,
         = quadra.nb.X,
         = quadra.nb.Y,
         = q.length.X,
         = q.length.X2,
         = q.length.Y,
         = q.length.Y2)

S_20_1 <- S_20_1 %>% 
  rename(Q_lengthX = Qlenght_X)

S_20_2 <- S_20_ %>% 
  rename(Q_lengthX = Qlenght_X)

S_25 <- S_25 %>% 
  select(Subplot, quadra.nb., q.length.X, q.length_1, q.length.Y, q.length_2) %>% 
  rename(Q_lengthX = q.length.X,
         = q.length_1,
         Q_lengthY = q.length.Y,
         = q.length_2)


```

