---
title: "Stakes_coordinates"
author: "Vincyane"
date: "`r Sys.Date()`"
output: html_document
---
# Packages
```{r Setup, include = F}
library(readr)
library(readxl)
library(tidyverse)
```

# Import data
Site : Paracou - P16
Subplots : 13-14-15-18-19-20-23-24-25
Méthode de mesure : laser

```{r, 13_18_23_24}
# 13, 18, 23, 24
S_13_18_23_24 <- read_delim("~/PhD/Inventories/Data/Piquets_ALT/Measures_piquets_Paracou13_18_23_24.csv", 
                            delim = ";", escape_double = FALSE, trim_ws = TRUE) # 401
names(S_13_18_23_24) # "plot" "quadra.nb.X" "quadra.nb.Y" "q.length.X"  "q.length.X2" "q.length.Y"  "q.length.Y2"
summary(S_13_18_23_24) # Min q.length.X : 1.0
S_13_18_23_24 <- S_13_18_23_24[complete.cases(S_13_18_23_24),] # enlever les lignes à NA -> 400
```
"plot" : subplot number
"quadra.nb.X" : number of column where the 10m x 10m quadrat is placed (0 - 9. 0 = sud; 9 = north).  
"quadra.nb.Y" : number of row where the 10m x 10m quadrat is placed (0 - 9. 0 = west; 9 = east).  
"q.length.X" : size of the southern side of the 10m x 10m quadrat (m).  
"q.length.X2" : size of the northern side of the 10m x 10m quadrat (m).  
"q.length.Y" : size of the western side of the 10m x 10m quadrat (m).  
"q.length.Y2" : size of the eastern side of the 10m x 10m quadrat (m).  

```{r}
S_14 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C14_measure.xlsx")
S_15 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C15_measure.xlsx")
S_19 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C19_measure.xlsx")
S_20_1 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C20_measure.xlsx") # 10/02/2025
S_20_2 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/ALT_Paracou_Quadrat_C20_fin20250130.xlsx") %>% # 06/02/2025
  select(Num, row_index, col_index, Qlenght_X, Q_lengthX1, Q_lengthY, Q_lengthY1, Subplot, Comments)
S_25 <- read_excel("~/PhD/Inventories/Data/Piquets_ALT/Measures_piquets_ParacouC25_T1_20230607ins.xlsx") %>%  select(Subplot, quadra.nb., quadra.n_1, q.length.X, q.length_1, q.length.Y, q.length_2) %>% 
  mutate(q.length.X = as.numeric(q.length.X),
         q.length_1 = as.numeric(q.length_1),
         q.length.Y = as.numeric(q.length.Y),
         q.length_2 = as.numeric(q.length_2)) %>% 
  unique()



all(names(S_14) == names(S_15))
all(names(S_14) == names(S_19))
names(S_20_1) == names(S_14)
names(S_20_2) # "left" "top""right" "bottom""row_index" "col_index" "Num"  "Qlenght_X" "Q_lengthX1" "Q_lengthY" "Q_lengthY1" "Subplot" "Comments"
names(S_25)

names(S_14) # "Num" "Q_lengthX" "Q_lengthX1" "Q_lengthY" "Q_lengthY1" "Subplot" "Comments"
summary(S_14) # Max length 100 !! c'est quoi Num; demander à Léa
summary(S_15)
summary(S_19)
summary(S_20_1)
summary(S_20_2) # Q_lengthY1 Max :109.000 !! demander à Léa;  diff entre S_20_1 et S_20_2
summary(S_25) # quadra.nb. tout à 0; quadra.n_1 max 3, ya pas tout je pense


nrow(S_14) == nrow(S_15)
nrow(S_19) == nrow(S_20_1)
nrow(S_20_2) == nrow(S_25)

```
"Num" :
"Q_lengthX" : size of the southern side of the 10m x 10m quadrat (m).
"Q_lengthX1" : size of the northern side of the 10m x 10m quadrat (m).
"Q_lengthY" : size of the western side of the 10m x 10m quadrat (m).
"Q_lengthY1" : size of the eastern side of the 10m x 10m quadrat (m).  
"Subplot" :  subplot number
"Comments" :
# Manage column names differences
```{r}
S_13_18_23_24 <- S_13_18_23_24 %>% 
  rename(Subplot = plot,
         # = quadra.nb.X,
         # = quadra.nb.Y,
         Q_lengthX = q.length.X,
         Q_lengthX1 = q.length.X2,
         Q_lengthY = q.length.Y,
         Q_lengthY1 = q.length.Y2)

S_20_1 <- S_20_1 %>% 
  rename(Q_lengthX = Qlenght_X)

S_20_2 <- S_20_ %>% 
  rename(quadra.nb.X = row_index, # demander à Léa
         quadra.nb.Y = col_index, # demander à Léa
         Q_lengthX = Qlenght_X)

S_25 <- S_25 %>% 
  rename(quadra.nb.X = quadra.nb., # demander à Léa
         quadra.nb.Y = quadra.n_1, # demander à Léa
         Q_lengthX = q.length.X,
         Q_lengthX1 = q.length_1, # demander à Léa
         Q_lengthY = q.length.Y,
         Q_lengthY1 = q.length_2) # demander à Léa

# Transformer Num en quadra.nb.X et quadra.nb.Y pour S_14, S_15, S_19, S_20_1 (demander à Léa)
# Num = 1 : quadra.nb.X = 0 et quadra.nb.Y = 0 ?
# Num = 2 : quadra.nb.X = 0 et quadra.nb.Y = 0 ?

```

# Bind
```{r}
Stakes <- bind_rows(S_13_18_23_24, S_20_1, S_20_2, S_25)
unique(Stakes$Comments)
```

# Eric's format
En plus de tes données, j’ai ajouté les « quadrats virtuels », c’est-à-dire les distances entre les piquets des bordures sud et ouest des parcelles pas encore mesurées : ce sont simplement les distances des bordures nord et est des placettes contigües.

plot	1- :ha subplot number
point_x	poi: nt number of the X-axis (0 to 10; left to right)
point_y	poi: nt number of the Y-axis (0 to 10; bottom to top)
d_left	dis: tance from the point to its left neighbor (point_x-1)
d_down	dis: tance from the point to its down neighbor (point_y-1)

```{r}

``Stakes %>% 
  rename(subplot = Subplot,
         point_x	= quadra.nb.X, # pas exactement il n'a pas 0-0 mais 0-10
         point_y	= quadra.nb.Y,
         d_left	=, 
         d_down =)`

