---
title: "Diagnose all DBH classes"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
---
```{r setup}
library(tidyverse)
library(cmdstanr)
library(bayesplot)

sp <- (read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/InterestSpecies.csv")[,3]) #[1:10]

Comb <- read.csv("D:/Mes Donnees/PhD/R_codes/PhD_cluster/Data/Combin_Sp_DBHclas.csv")

unique(Comb$DBH_classes)
# 75
# mettre le 2 avec la correc bota


strategy <- read_delim("~/PhD/Inventories/Data/Understory/Paracou/Species_strategy.csv") %>% 
  select(ScientificName, Strategy_VB) %>%
  rename(Strategy = Strategy_VB) %>% 
  unique()

PATH <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots/1-3/"
```

# Fits
```{r fit}
fits <- list()
for(S in sp){
  # Filter species with only one DBH class
  if(length(list.files(file.path("Chains", "Hybrid_eigenvectors", S)))>1){
    
    for(D in c("1-3","3-10","10-25","-25")){
      # for(D in c("1-5","5-10","10-20","-20")){
      tryCatch({
        chain_path <- file.path("Chains", "Hybrid_eigenvectors", S, D) #  Hybrid_eigenvectors Hybrid_no_iota
        
        if(D=="-25") D <- ">25"
        # if(D=="-20") D <- v">20"
        
        fits[[S]][[D]] <- as_cmdstan_fit(list.files(chain_path,
                                                    full.names = TRUE))},
        error=function(e){cat("ERROR :",S, D, conditionMessage(e), "\n")}
      )
    }
  }}
# names(fits[["Symphonia_sp.1"]])

names(fits)[names(fits) == "-25"] <- ">25"
# names(fits)[names(fits) == "-20"] <- ">20"


# 70 sp with more than 1 DBH class
```

# Divergence
```{r divergence}
d <- list()
for(S in names(fits)){
  for(D in c("1-3","3-10","10-25",">25")){
    # for(D in c("1-5","5-10","10-20",">20")){
    
    if(D %in% names(fits[[S]])){
      
      d[[S]][[D]] <- data.frame(Species = S,
                                DBH = D,
                                Chain = c(1:4),
                                Divergences = fits[[S]][[D]]$diagnostic_summary("divergences") # divergences per chain
      )
    }}}
Div <- list_rbind(list_flatten(d)) %>% 
  bind_rows() %>% 
  rename(Divergences = num_divergent) %>% 
  group_by(Species) %>% 
  mutate(`%divergence` = sum(Divergences)/4000*100)

```

# Rhat
R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1.01 (Vehtari et al., 2021). 
R-hat which is the maximum of rank normalized split-R-hat and rank normalized folded-split-R-hat, which works for thick tailed distributions and is sensitive also to differences in scale.
```{r Rhat}
d <- list()
for(S in names(fits)){
  for(D in c("1-3","3-10","10-25",">25")){
    # for(D in c("1-5","5-10","10-20",">20")){
    
    if(D %in% names(fits[[S]])){
      
      d[[S]][[D]] <- fits[[S]][[D]]$summary(c("alpha", "beta1", "beta2_p", "tau")) %>% 
        mutate(Species = S) %>% 
        mutate(DBH = D) %>% 
        select(Species, DBH, variable, rhat) 
    }}}
Rhat_table <- list_rbind(list_flatten(d))


# bayesplot::rhat(fits[[1]],pars="beta2_p")
# posterior::rhat(extract_variable_matrix(fits[[1]], "beta2_p"))

all(Rhat_table$rhat < 1.02)

badRhat <- Rhat_table %>% filter(rhat > 1.02) %>%
  left_join(Div %>% select(Species, `%divergence`) %>% unique(), by="Species")

length(unique(badRhat$Species))
# (40/70)*100 = 53% for 1-5 classes (sans filtre pour le nombre d'individu)
# (46/70)*100 = 66% for 1-3 classes
# (26/70)*100 = 37% for 1-3 classes with autocor

```


# Chains
```{r}
name <- 
  # names(fits)[names(fits) %in% c("Anaxagorea_dolichocarpa", "Dicorynia_guianensis", "Iryanthera_hostmannii", "Protium_opacum", "Talisia_guianensis", "Talisia_sylvatica", "Eugenia_coffeifolia", "Ryania_speciosa", "Inga_loubryana", "Eschweilera_sagotiana", "Hirtella_bicornis", "Tabernaemontana_macrocalyx", "Talisia_mollis", "Eugenia_brownsbergii", "Duroia_longiflora", "Mouriri_sagotiana", "Ixora_ferrea")]
  unique(badRhat$Species)
# sort(names(fits))

for(S in name){
  for(D in c("1-3","3-10","10-25",">25")){
    # for(D in c("1-5","5-10","10-20",">20")){
    # if(D %in% names(fits[[S]])){
    if(any(D %in% badRhat[badRhat$Species==S,]$DBH)){
      # S = "Talisia_sylvatica"
      # D = "5-10"
      print(
        fits[[S]][[D]]$draws(
          c("lp__", "alpha", "beta1", "beta2_p", "O", "tau")
          # c("lp__", "a", "O", "gamma", "tau")
        ) %>% 
          mcmc_trace() +
          # mcmc_dens() +
          ggtitle(paste(S, D))
      )
    }
  }}
```

# Filter species with bad Rhat
```{r}
fits <- fits[!names(fits) %in% name] # 44 sp
```

# Prepare data for plots
```{r}
d <- list()
# Plot data
for(S in names(fits)){
  for(D in c("1-3","3-10","10-25",">25")){
    if(D %in% names(fits[[S]])){
      
      d[[S]][[D]] <- mcmc_intervals_data(fits[[S]][[D]]$draws(c("a", "O", "tau")),
                                         prob = 0.5, prob_outer = 0.95, point_est = "median") %>% 
        mutate(Species = S) %>% 
        mutate(DBH = D)
      
    }}}
datap <- list_rbind(list_flatten(d))

# Light effect (not flat)
O_sg <- datap %>% 
  # filter(parameter=="O"& incertitude > 4) %>%
  filter(parameter=="a"& m > -0.02) %>% # no light effect
  select(Species, DBH) %>% 
  mutate(parameter="O") %>% 
  mutate(Sg_effect = F)

datap <- datap %>% 
  left_join(O_sg, by=c("Species", "DBH", "parameter")) %>% 
  # tau significantly different of 0
  mutate(Sg_effect = ifelse(parameter=="tau" & ll<0 & 0<hh, F, Sg_effect))  %>% 
  mutate(Sg_effect = ifelse(is.na(Sg_effect), T, Sg_effect))  %>% 
  mutate(DBH = factor(DBH, levels=c("1-3","3-10","10-25",">25"))) %>% 
  left_join(strategy, by = c("Species" = "ScientificName")) %>% 
  mutate(Strategy = factor(Strategy, levels=c("Canopy species","Understory species"))) %>% 
  arrange(Species) %>% arrange(Strategy) %>% 
  mutate(Species=factor(Species, levels=unique(Species))) %>% 
  mutate(Topo= case_when(
    parameter=="tau" & m > 0 ~ "More flood tolerant",
    parameter=="tau" & m < 0 ~ "Less flood tolerant")) %>% 
  mutate(Topo = ifelse(parameter=="tau" & ll<0 & 0<hh, "Generalist", Topo))

# -a
datap <- datap %>%
  mutate_at(vars(c("ll", "l", "m", "h","hh")), ~ ifelse(parameter=="a", (-.), .))

ord_sp <- unique(datap$Species)

species_to_number <- setNames(seq_along(ord_sp), ord_sp)

numeric_vector <- species_to_number[datap$Species]

datap$sp_code <- numeric_vector

# ll : outer start
# l : iner start
# m : point_est value
# h : iner end
# hh : outer end

```

# Posteriors
ajouter le alpha sur O
fixer y pour toutes les pages
```{r}
# mcmc_intervals() code 
# trace(bayesplot::mcmc_intervals, edit=T)

for(P in c('O', 'a', 'tau')){
  i = 10 # nbr of sp per page
  plist <- vector('list', round(length(names(fits))/i))
  
  for(p in 1:round(length(names(fits))/i)){
    # p1 to 8 for 10 sp each
    # message(s)
    plist[[p]] <- local({
      # p <- 7
      if (p==1) {start = 1
      }else{start = i*(p-1) + 1}
      stop = start + (i-1)
      
      # print(p) ;  print(P)
      
      datas <- datap %>% 
        filter(parameter==P) %>%
        filter(sp_code >= start & sp_code <= stop)
      
      ggplot(datas, aes(x= DBH, y = m, group=1)) +
        theme_minimal() + 
        geom_segment(aes(y=ll, yend=hh, x=DBH, xend=DBH,
                         color = "95%"), size = 0.5) +
        geom_segment(aes(y=l, yend=h, x=DBH, xend=DBH,
                         color = "50%"), size = 1) +
        scale_color_manual(name = "Credible interval", 
                           values = c("95%" = "#E7B800",
                                      "50%"= "#FC4E07")) + 
        {if(P=='a') geom_point() } + # median
        {if(P!='a') geom_point(aes(alpha = Sg_effect)) } +
        geom_line() +
        {if(P=='tau') geom_hline(yintercept=0, linetype = "dashed", col = "red") } +
        scale_x_discrete(position = "top") +
        
        scale_y_continuous(position = "right") +
        # {if(P=='O') scale_y_continuous(position = "right") } +
        # {if(P=='tau') scale_y_continuous(position = "right", limits = c(-20, NA))  } +
        # {if(P=='a')  scale_y_continuous(position = "right", limits = c(-3, NA)) } +
        
        # , limits = c(min(datas$ll), max(datas$hh))
        labs(x= 'DBH (cm)', y= P) +
        facet_grid(Species ~ ., switch = "y") +
        theme(strip.text.y.left = element_text(angle = 0))
    })
  }
  
  plots <- gridExtra::marrangeGrob(plist, nrow = 1, ncol = 1)
  # plots
  
  
  ggsave(paste(P,"_posteriors_stages_autocor.pdf", sep=''), path = PATH, # _sg
         plot = plots, width = 25, height = 26, units = "cm", dpi=800, bg="white")
}
```

# Parameters accross onto
```{r}
for(S in c('Understory species', 'Canopy species')){
  # S='Canopy species'
  for(P in c('a', 'O', 'tau')){
    # P='a'
    
    datap %>% 
      filter(parameter==P) %>% 
      filter(Strategy==S) %>% 
      ggplot(aes(x= DBH, y = m, col = Species, group = Species, alpha= Sg_effect)) + # 
      theme_minimal() + 
      {if(P!='tau') list(geom_point(), geom_line(), guides(col = guide_legend(ncol = 2)) ) } + # median
      {if(P=='tau') list(geom_point(aes(col=Topo)), geom_line(aes(col=Topo)) ) } + # , scale_y_continuous(limits = c(-20, NA))
      # geom_segment(aes(y=l, yend=h, x=DBH, xend=DBH), size = .5) + # 5O% CI
      
      {if(P=='O')  scale_y_continuous(limits = c(NA,0), n.breaks = 10, labels = ~round(exp(.)*100,2)) } +
      scale_alpha_manual(values = c("FALSE" = .2, "TRUE"= 1)) +
      # guides(color = F) +
      {if(P=='a')  list(guides(alpha = F)) } + # , scale_y_continuous(limits = c(-3, NA))
      
      {if(P=='tau') geom_hline(yintercept=0, linetype = "dashed", col = "red") } +
      labs(title=S, y=P)
    
    ggsave(paste(S,P,"accross_onto_autocor.png", sep='_'), path = PATH,
           width = 25, height = 26, units = "cm", dpi=800, bg="white")
  }
}
```

```{r}
datap %>% 
  select(Species, DBH, Topo) %>% 
  na.omit() %>% 
  # summarise(N = n()) %>% 
  ggplot(aes(x= DBH, fill= Topo)) +
  theme_minimal() + 
  geom_histogram(stat="count") +
  labs(y = 'Species number', fill = "Topographic sensitivity")
```

# Boxplot parameters accross onto
```{r}
for(S in c('Understory species', 'Canopy species')){
  for(P in c('a', 'O', 'tau')){
    print(
      datap %>% 
        filter(parameter==P) %>% 
        filter(Strategy==S) %>% 
        ggplot(aes(x= DBH, y = m)) +
        theme_minimal() + 
        geom_boxplot() +
        {if(P=='tau') scale_y_continuous(limits = c(-20, NA))  } +
        {if(P=='a')  scale_y_continuous(limits = c(-2, NA)) } +
        
        labs(title=S, y=P)
    )
    # ggsave(paste(S,"a_boxplot_accross_onto_sg.png", sep='_'), path = PATH,
    #        width = 25, height = 26, units = "cm", dpi=800, bg="white")
  }
}
```

# Predictions
il faut l'environnement de chaque classe onto...
1 plot par sp et une couleur par stade
```{r}
Envpred <- read_csv("D:/Mes Donnees/PhD/R_codes/PhD/Modelisation/Realdata/Envrmt_prediction.csv") %>% 
  mutate(DBH_classes = recode(DBH_classes, "-25" = ">25"))
```


```{r predictions}
Sys.time()

datalist <- list()
Olist <- list()
plist <- list()

for(s in names(fits)){
  plist[[s]] <- local({
    # s= "Lecythis_persistens"
    for(d in c("1-3","3-10","10-25",">25")){
      # d= ">25"
      if(d %in% names(fits[[s]])){
        
        datalist[[d]] <- fits[[s]][[d]]$summary("p") %>%  # p posterior
          # prediction environment 
          rowwise() %>% 
          mutate(DBH = d) %>% 
          mutate(Light = as.numeric(unlist(str_extract_all(variable, "\\d+"))[1])) %>% 
          # mutate(Topo = as.numeric(unlist(str_extract_all(variable, "\\d+"))[2])) %>% 
          mutate(logTransmittance = Envpred[Envpred$DBH_classes==d,]$Lightp[Light]) 
        # mutate(logTWI = Envpred[Envpred$DBH_classes==d,]$Topographyp[Topo])
        
        Olist[[d]] <- data.frame(
          DBH = d,
          O = fits[[s]][[d]]$summary("O", median)$median,
          OQ5 = fits[[s]][[d]]$summary("O")$q5, # Opt Q5%
          OQ95 =fits[[s]][[d]]$summary("O")$q95
        )
        
      }
    }
    DATA <- list_rbind(list_flatten(datalist)) %>% 
      mutate(DBH = factor(DBH, levels=c("1-3","3-10","10-25",">25")))
    ODATA <- list_rbind(list_flatten(Olist)) %>% 
      mutate(DBH = factor(DBH, levels=c("1-3","3-10","10-25",">25")))
    
    DATA %>% 
      ggplot(aes(col = DBH, fill=DBH)) +
      theme_minimal() +
      geom_line(aes(x = logTransmittance, y = median), size=0.7) + # p posterior median
      geom_ribbon(aes(x = logTransmittance, ymin = q5, ymax = q95),
                  alpha = .1, col=NA, linewidth=0.5) + # P quantiles
      geom_vline(data = ODATA, aes(xintercept  = O, col= DBH), linewidth=0.8) + # Opt median
      
      geom_rect(data = ODATA, aes(xmin = OQ5, xmax = OQ95, # O quantiles
                                  ymin=0, ymax=Inf), fill=NA, linetype="dotted") +
      
      # geom_vline(data = ODATA, aes(xintercept  = OQ5, col= DBH),  linetype="dashed") + # Opt Q5%
      # geom_vline(data = ODATA, aes(xintercept  = OQ95, col= DBH),  linetype="dashed") + # Opt Q95%
      scale_color_manual(values = c("1-3" = "#00AFBB",
                                    "3-10"= "#CC79A7",
                                    "10-25"= "#E7B800",
                                    ">25" = "#FC4E07"),
                         breaks=c("1-3","3-10","10-25",">25")) + 
      
      scale_fill_manual(values = c("1-3" = "#00AFBB",
                                   "3-10"= "#CC79A7",
                                   "10-25"= "#E7B800",
                                   ">25" = "#FC4E07"),
                        breaks=c("1-3","3-10","10-25",">25")) + 
      
      scale_x_continuous(name = "Light (% transmittance)",
                         n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
      ggtitle(paste(s,"- at logTWI optimum")) +
      labs(y="Presence probabiblity median", col='DBH (cm)', fill='DBH (cm)')
    
  })
}


plots <- gridExtra::marrangeGrob(plist, nrow = 3, ncol = 3) # 75
# plots

ggsave("Allsp_Light_niche_atTWIoptimum_stages_autocor.pdf",
       path = PATH,
       plots, width = 15, height = 10)

Sys.time() # 4 min
```

# Biplot O vs a
```{r}
databp <- datap %>% 
  select(-c(outer_width, inner_width, point_est, Sg_effect)) %>% 
  rename(Q5 = ll, Q95 = hh, median = m, Q50_0 = l, Q50_1 = h) %>% 
  filter(parameter %in% c("O", "a")) %>% 
  pivot_wider(names_from = parameter, values_from = c(Q5, Q50_0, median, Q50_1, Q95),
              names_glue = "{parameter}_{.value}") %>% 
  mutate(Sg_effect = ifelse(a_median > -0.02, F, T))  %>% # no light effect
  mutate(a_median = -a_median)

# databp <- databp %>%
#   mutate_at(vars(c(a_Q5, a_Q50_0, a_median, a_Q50_1, a_Q95)), ~ (-.))

```

```{r}
databp %>% 
  # filter(Species %in% c("Ruizterania_albiflora", "Pouteria_gonggrijpii", "Trichilia_schomburgkii", "Virola_michelii", "Lecythis_corrugata", "Symphonia_sp.1", "Dicorynia_guianensis", "Anaxagorea_dolichocarpa", "Vouacapoua_americana")) %>% 
  ggplot(aes(col = DBH, group = Species, shape = Strategy)) +
  theme_minimal() + 
  # geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin= xstart,
  #                           xmax=xend, alpha=light), fill="grey") +
  
  # geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin= min(databp$a_Q5),
  #                           xmax=xend, alpha=light), fill="grey") +
  # scale_alpha_manual(name = "Light preference", values=c(.1,.5)) +
  
  # 95% credible interval
  # geom_segment(aes(x=a_Q5, xend=a_Q95, y=O_median, yend=O_median,
  #                  col = "95 and 50% CI"), size = .5) + # tau
  # geom_segment(aes(y=O_Q5, yend=O_Q95, x=a_median, xend=a_median,
  #                  col = "95 and 50% CI"), size = .5) + # O
  # geom_segment(aes(x=a_Q50_0, xend=a_Q50_1, y=O_median, yend=O_median,
  #                  col = "95 and 50% CI"), size = .6) + # tau
  # geom_segment(aes(y=O_Q50_0, yend=O_Q50_1, x=a_median, xend=a_median,
  #                  col = "95 and 50% CI"), size = .6) + # O
  
  # geom_line(aes(x= a_median, y = O_median), col = "grey") +
  geom_point(aes(x= a_median, y = O_median), size =3) + # median
  # scale_shape_manual(name="Light effect",
  #                    values=c("FALSE" = 1, "TRUE" = 19),
  #                    breaks = c("TRUE","FALSE")) +
  # scale_color_manual(name="", values = c("95 and 50% CI" = "darkgrey")) +
  scale_color_manual(values = c("1-3" = "#00AFBB",
                                "3-10"= "#CC79A7",
                                "10-25"= "#E7B800",
                                ">25" = "#FC4E07"),
                     breaks=c("1-3","3-10","10-25",">25")) + 
  
  
  ggrepel::geom_text_repel(aes(x= a_median, y = O_median, label = Species), col = "black", size = 2.5) +
  
  # stat_ellipse(aes(x= a_median, y = O_median, color = as.factor(cluster)),
  #              type = "t",geom = "polygon",alpha = 0.4) +
  
  scale_y_continuous(n.breaks = 10, labels = ~round(exp(.)*100,2)) + # delog
  scale_x_continuous(trans="log") +
  theme(axis.title = ggtext::element_markdown(),
        legend.title = ggtext::element_markdown(),
        legend.position="bottom") +
  labs(x= "Niche width (log(-*a*))",
       y= "Light optimum (*O*) (% transmittance)")

# alpha sur les postériors imprécis

ggsave("O_a_biplot_onto_autocor.png", path = PATH, 
       width = 32, height = 20, units = "cm", dpi=800, bg="white")

```


# ACP O, a, incertitude O, incertitude a