---
title: "Diagnose all"
date: "`r Sys.Date()`"
format: html
self-contained: true
theme: cosmo
editor: source
code-fold: true
---

```{r setup}
library(tidyverse)
library(cmdstanr)
library(bayesplot)

sp <- (read.csv("D:/Mes Donnees/PhD/Inventories/Data/Understory/Paracou/InterestSpecies.csv")[,3]) # 75
# metrre le 2 avec la correc bota

# sp <- c("Anaxagorea_dolichocarpa", "Tabernaemontana_macrocalyx", "Eperua_falcata", "Dicorynia_guianensis", "Paypayrola_hulkiana", "Symphonia_sp.1", "Pogonophora_schomburgkiana")

# sp <- "Iryanthera_hostmannii"

```

```{r fit}
fits <- list()
for(S in sp){
  tryCatch({
    chain_path <- file.path("Chains", "Hybrid", S)
    fits[[S]] <- as_cmdstan_fit(list.files(chain_path,
                                           full.names = TRUE))},
    error=function(e){cat("ERROR :",S, conditionMessage(e), "\n")}
  )
}
# names(fits[["Symphonia_sp.1"]])
```

```{r divergence}
Div <- lapply(names(fits), function(x)
  data.frame(Species = x,
             Chain = c(1:4),
             Divergences = fits[[x]]$diagnostic_summary("divergences") # divergences per chain
  ) 
) %>% 
  bind_rows() %>% 
  rename(Divergences = num_divergent) %>% 
  group_by(Species) %>% 
  mutate(`%divergence` = sum(Divergences)/4000*100)

```

# Rhat
R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1.01 (Vehtari et al., 2021). 
R-hat which is the maximum of rank normalized split-R-hat and rank normalized folded-split-R-hat, which works for thick tailed distributions and is sensitive also to differences in scale.
```{r Rhat}
Rhat_table <- lapply(names(fits), function(x)
  fits[[x]]$summary(c("alpha", "beta1", "beta2_p", "tau", "iota")) %>% 
    mutate(Species = x) %>% 
    select(Species, variable, rhat) 
) %>% 
  bind_rows()

# bayesplot::rhat(fits[[1]],pars="beta2_p")
# posterior::rhat(extract_variable_matrix(fits[[1]], "beta2_p"))

all(Rhat_table$rhat < 1.01)

badRhat <- Rhat_table %>% filter(rhat > 1.01) %>%
  left_join(Div %>% select(Species, `%divergence`) %>% unique(), by="Species")
```

```{r}
name <- 
  # "Iryanthera_hostmannii"
  # c("Guarea_costata", "Virola_michelii",  "Cordia_nodosa",
#   #   "Paypayrola_hulkiana", "Pouteria_guianensis") # good now
  
# names(fits)[names(fits) %in% c("Anaxagorea_dolichocarpa", "Dicorynia_guianensis", "Iryanthera_hostmannii", "Protium_opacum", "Talisia_guianensis", "Talisia_sylvatica", "Eugenia_coffeifolia", "Ryania_speciosa", "Inga_loubryana", "Eschweilera_sagotiana", "Hirtella_bicornis", "Tabernaemontana_macrocalyx", "Talisia_mollis", "Eugenia_brownsbergii", "Duroia_longiflora", "Mouriri_sagotiana", "Ixora_ferrea")]
# unique(badRhat$Species) 
sort(names(fits))

lapply(name , function(x)
  fits[[x]]$draws(
    c("lp__", "alpha", "beta1", "beta2_p", "O", "iota")
    # c("lp__", "a", "O", "gamma", "tau", "iota")
    ) %>% 
    mcmc_trace() +
    # mcmc_dens() +
    ggtitle(x)
)
```

```{r}
Lightp = seq(-7.308724, 8.941571e-08, length.out = 50)
Topographyp = 1.764199
dbhs <- c(2.5, 7.5, 15, 25, 35)
post <- as.data.frame(fits[["Iryanthera_hostmannii"]]$draws(c("a", "O", "iota", "beta1", "alpha", "p"))) # 2000 rows 4 cols per param

post_long <- post %>%
  pivot_longer(
    cols = everything(),
    names_to = c("chain", "variable"),
    names_sep = "\\.",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  ) %>%
  unnest() %>% 
  pivot_longer(
    cols = starts_with("P["),
    values_to = "p"
  ) %>% 
  mutate(name = str_remove(name, "p")) %>% 
  rowwise() %>% 
  mutate(Env = as.numeric(unlist(str_extract_all(name, "\\d+"))[1])) %>% 
  mutate(dbh = as.numeric(unlist(str_extract_all(name, "\\d+"))[2])) %>% 
  mutate(Environmentp = Lightp[Env]) %>% # 50
  mutate(DBHp = dbhs[dbh]) # 5

# post_long <- post_long %>% 
#   group_by(chain) %>% 
#   summarise(~mean)
# mutate(p = brms::inv_logit(a*(Lightp - (O + iota*DBHp[i]))^2 + gamma + tau*Topographyp))
```

```{r}
post_long %>% 
  filter(DBHp %in% c(2.5, 35)) %>% 
  ggplot(aes(x = Environmentp, y = p, col = chain, linetype = as.factor(DBHp))) +
  theme_minimal() +
  geom_smooth()

post_long %>% 
  filter(DBHp == 2.5) %>% 
  ggplot(aes(x = a, y = iota, col = chain)) +
  theme_minimal() +
  geom_line()

post_long %>% 
  filter(DBHp == 2.5) %>% 
  ggplot(aes(x = alpha, y = iota, col = chain)) +
  theme_minimal() +
  geom_line()

post_long %>% 
  filter(DBHp == 2.5) %>% 
  ggplot(aes(x = beta1, y = iota, col = chain)) +
  theme_minimal() +
  geom_line()

```


# Posteriors
Plot central (quantile-based) posterior interval estimates from MCMC draws
```{r}
# Plot data
datap <- lapply(names(fits), function(x)
  mcmc_intervals_data(fits[[x]]$draws(c("a", "O", "gamma", "tau", "iota")),
                      prob = 0.5, prob_outer = 0.95, point_est = "median") %>% 
    mutate(Species = x)
)  %>% 
  bind_rows()
# ll : outer start
# l : iner start
# m : point_est value
# h : iner end
# hh : outer end

# mcmc_intervals() code 
# trace(bayesplot::mcmc_intervals, edit=T)

# 0 line
x_lim <- range(c(datap$ll, datap$hh))
x_range <- diff(x_lim)
x_lim[1] <- x_lim[1] - 0.05 * x_range
x_lim[2] <- x_lim[2] + 0.05 * x_range

# 0 line
layer_vertical_line <- if (0 > x_lim[1] && 0 < x_lim[2]) {
  vline_0(color = "darkred", linewidth = 0.5, linetype = "dashed")
}

ggplot(datap, aes(y = Species)) +
  theme_minimal() + 
  # xlim(range(c(datap$ll, datap$hh))) +
  layer_vertical_line +
  geom_segment(aes(x=ll, xend=hh, y=Species, yend=Species,
                   color = "95%"), size = 0.5) +
  geom_segment(aes(x=l, xend=h, y=Species, yend=Species,
                   color = "50%"), size = 1) +
  geom_point(aes(x= m)) + # median
  scale_color_manual(name = "Credible interval", 
                     values = c("95%" = "#E7B800",
                                "50%"= "#FC4E07")) + 
  theme(axis.title.x =element_blank()) +
  facet_grid(~parameter, scale= "free")


ggsave("Light_topo_onto_posteriors_iotaprior.pdf", path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
       width = 25, height = 26, units = "cm", dpi=800, bg="white")

# O et tau iront dans le biplot
# les paramètres dont on veut la significativité : iota
# a on veut sa valeur absolue pour juger la largeur de niche
# gamma on s'en fout ?
```


```{r}
fits[[1]]$draws(c("a", "O", "gamma", "tau")) %>% mcmc_areas_ridges()
```

# Biplot
O in y, tau in x
```{r}
datap %>% 
  select(-c(outer_width, inner_width, l, h, point_est)) %>% 
  rename(Q5 = ll, Q95 = hh, median = m) %>% 
  filter(parameter %in% c("O", "tau")) %>% 
  pivot_wider(names_from = parameter, values_from = c(Q5, median, Q95),
              names_glue = "{parameter}_{.value}") %>% 
  
  ggplot(aes(x= tau_median, y = O_median)) +
  theme_minimal() + 
  # 95% credible interval
  geom_segment(aes(x=tau_Q5, xend=tau_Q95, y=O_median, yend=O_median), size = 0.5, col = "grey") + # tau
  geom_segment(aes(y=O_Q5, yend=O_Q95, x=tau_median, xend=tau_median), size = 0.5, col = "grey") + # O
  geom_point() + # median
  ggrepel::geom_text_repel(aes(x= tau_median, y = O_median, label = Species), col = "black", size = 3) +
  labs(x= "TWI effect (tau)", y= "Light optimum (O)")


```


# Predictions
```{r predictions}
Sys.time()

plist <- list()

for(s in sp){
  plist[[s]] <- local({
    
    g <- lapply(names(fits[[s]]), function(v){
      # v = "TWI"
      
      DATA <- fits[[s]][[v]]$summary("p") # p posterior
      # prediction environment 
    })
    
  })
}

plots <- gridExtra::marrangeGrob(plist, nrow = 3, ncol = 3) # 75
# plots

# ggsave("Allsp_Light_TWI_niche.pdf",
#        path = "//amap-data.cirad.fr/work/users/VincyaneBadouard/Modelisation/Plots",
#        plots, width = 15, height = 10)

Sys.time()
```


