---
title: "Light maps"
author: "Vincyane Badouard"
date: "`r Sys.Date()`"
output: pdf_document
---

# Packages
```{r, include = F}
library(data.table)
library(tidyverse)
library(sf)
library(terra)
```

# Spatial objects
```{r}
buffer <- st_as_sf(vect("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/Plot16_4ha_buffer.shp")) # interest zone with buffer
zone <- st_as_sf(vect("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2022/Parcelles_Understory.shp")) # 4 carrés
Understory_stakes <- read.csv("D:/Mes Donnees/PhD/SIG_data/Understory/Understory_stakes_theoretical_UTM.csv")


```

# Process raw data from AMAPVox
```{r}
path <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_annual.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_Dry.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_Wet.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_buffer_LowAlt_Light_annual.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2022/P16_2022_4ha_buffer_Light_annual.txt"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ.txt"
```

```{r}
Light <- fread(path, skip=10) # skip the 10 1st lines of header
Light <- Light[Y != "MEAN"] # enlever les 2 lignes mean
Light <- Light[, !c("V5"), with = FALSE] # remove empty column
Light[, `:=`(X = as.numeric(X) , Y = as.numeric(Y))] # all as numeric
Light <- Light[!is.na(X)] 
```
## Plot
```{r}
# Remove edges
# Light <- Light[X<90 & X> -90 & Y<90 & Y> -90]


# Plot
lattice::levelplot(`Period 1` ~ X*Y, data = Light) # Transmittance
lattice::levelplot(Z~X*Y, data = Light) # Z = Altitude

ggplot() + 
  geom_raster(data = Light, aes(X, Y, fill = `Period 1`)) + 
  scale_fill_gradientn(name = "Light intensity (Transmittance)",
                       colors = terrain.colors(30, rev=T),
                       na.value="white") +
  ggtitle("Paracou P16 4ha - Light map 1m") +  
  coord_sf()
```

## Transformation des coordonnées relatives en UTM
On peut retrouver les cooordonnées UTM avec la VOP
```{r}
# Lecture matrice de transformation
path <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/VOP_P16_9ha.txt"

VOP <- as.matrix(read.table(path))

# Apply inverse of VOP matrix to convert back to UTM
xyz <- tcrossprod(as.matrix(Light[, .(X, Y, Z, c=1)]), solve(VOP))
Light[, `:=`(Xutm = xyz[, 1], Yutm = xyz[, 2], Zutm = xyz[, 3])]

Light <- unique(Light)

ggplot() + 
  geom_point(data = Light, size = 1.28,
             aes(Xutm, Yutm, color = `Period 1`)) + 
  scale_color_gradientn(name = "Light intensity (Transmittance)",
                        colors = terrain.colors(30, rev=T),
                        na.value="white") +
  geom_sf(data = sf::st_cast(buffer, "LINESTRING")) + # buffer (to remove after)
  geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
  ggtitle("Paracou P16 4ha - Light map 1m") +  
  coord_sf()
```

## Crop on the interst zone
```{r}
Lightsf <- st_as_sf(Light, coords = c('Xutm','Yutm'))
Lightsf <- st_set_crs(Lightsf, crs(zone))
Lightcrop <- st_intersection(Lightsf, zone)
XY <- st_coordinates(Lightcrop) # stock coordinates
colnames(XY) <- c("Xutm", "Yutm") 
st_geometry(Lightcrop) <- NULL # no more sf object
Lightcrop <- cbind(Lightcrop, XY) # bind XY coord


ggplot() + 
  geom_point(data = Lightcrop, size = 1.28,
             aes(Xutm, Yutm, color = `Period.1`)) + 
  scale_color_gradientn(name = "Light intensity (Transmittance)",
                        colors = terrain.colors(30, rev=T),
                        na.value="white") +
  # geom_sf(data = sf::st_cast(buffer, "LINESTRING")) + # buffer 
  geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
  ggtitle("Paracou P16 4ha - ALS High Altitude - Dry season - Light map 1m") +  
  coord_sf()
```

## Write data cropped and in UTM
```{r}
fwrite(Light, , row.names = F, col.names = T)
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.txt"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_UTM.txt"


write.csv(Light, )
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.csv"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_UTM.csv"

```

# Load data pré-processed
```{r}
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.txt"

Light_HighAlt <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.txt")

Light_LowAlt <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.txt")
```

# Maps

#  Light differences
leftjoin by coordinates 
difference between `Period.1` of the 2 datasets
```{r}
names(Light_HighAlt) # "X","Y","Z","Period.1","Zutm","SubPlotCon","Forest","Plot","SubPlot","Projet","Protocole","Xutm","Yutm"

Light <- merge(Light_HighAlt, Light_LowAlt, by=c("Xutm","Yutm","Zutm", "X","Y","Z", "Forest","Plot","SubPlot","Projet","Protocole","SubPlotCon"), all=TRUE, suffixes = c("_HighAlt", "_LowAlt")) # merge datasets

Light[,Lightdiff:=`Period.1_LowAlt`-`Period.1_HighAlt`] # compute differences (4-3=1, le rslt informe sur la valeur à laquelle on soustrait)
```

## Maps
```{r}
ggplot() + 
  geom_point(data = Light, size = 1.28,
             aes(Xutm, Yutm, color = `Period.1_LowAlt`)) + 
  scale_color_gradientn(name = "Light intensity (Transmittance)",
                        colors = terrain.colors(30, rev=T),
                        na.value="white") +
  geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
  geom_point(data = Understory_stakes, size = 0.5,
             aes(Xutm, Yutm)) +
  ggtitle("Paracou P16 4ha - ALS Low Altitude - Light map 1m") +  
  coord_sf()
```

```{r}
hist(Light$Lightdiff, breaks = 15)
Min <- min(Light$Lightdiff)
Max <- max(Light$Lightdiff)

ggplot() + 
  geom_point(data = Light, size = 1.28,
             aes(Xutm, Yutm, color = Lightdiff)) + 
  # scale_colour_gradient2(name = "Light intensity (Transmittance) difference", low="#00AFBB", high="#FC4E07", mid = "white", midpoint = 0)+
  scale_colour_steps2(name = "Light intensity (Transmittance) difference",
                      low="#000066", high="red", mid = "white", midpoint = 0.00,
                      limits = c(Min, Max),
                      breaks = c(Min, -0.06, -0.04, -0.02, -0.01, 0.01, 0.02, 0.04, 0.06, Max), # c(Min, -0.08, -0.02, 0.02, 0.06, Max) ; seq(Min, Max, by = 0.02)
                      labels = as.character(round(c(Min, -0.06, -0.04, -0.02, -0.01, 0.01, 0.02, 0.04, 0.06, Max), digits = 2))) +
  # geom_sf(data = sf::st_cast(buffer, "LINESTRING")) + # buffer 
  geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
  ggtitle("Paracou P16 4ha - ALS High vs Low Altitude 2023 - Light map difference (1m)") +  
  coord_sf()

ggplot() + 
  geom_point(data = Light, size = 1.28,
             aes(Xutm, Yutm, color = Lightdiff)) + 
  # scale_colour_gradient2(name = "Light intensity (Transmittance) difference", low="#00AFBB", high="#FC4E07", mid = "white", midpoint = 0)+
  scale_colour_steps2(name = "Light intensity (Transmittance) difference",
                      low="#000066", high="red", mid = "white", midpoint = 0.00,
                      limits = c(Min, Max),
                      breaks = seq(-0.1, 0.1, by = 0.025) # c(Min, -0.08, -0.02, 0.02, 0.06, Max) ; seq(Min, Max, by = 0.02)
  ) +
  # geom_sf(data = sf::st_cast(buffer, "LINESTRING")) + # buffer 
  geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
  ggtitle("Paracou P16 4ha - ALS High vs Low Altitude 2023 - Light map difference (1m)") +  
  coord_sf()
```

# Light profiles
```{r}
Lightgrid_HighAlt <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_UTM.txt")

Lightgrid_LowAlt <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_UTM.txt")

```

## Plot
```{r}
names(Lightgrid_HighAlt) # "X","Y","Z",  "Period 1",  "Xutm","Yutm","Zutm"    

dataHigh <- Lightgrid_HighAlt %>% 
  mutate(Zutm_disc = as.integer(Zutm)) %>% 
  group_by(Zutm_disc) %>% 
  mutate(MeanTransmittance = mean(`Period 1`)) %>% 
  mutate(MinTransmittance = min(`Period 1`)) %>% 
  mutate(MaxTransmittance = max(`Period 1`)) %>%
  mutate(SdTransmittance = sd(`Period 1`)) %>% 
  ungroup() %>% 
  select(Zutm_disc, MeanTransmittance, MinTransmittance, MaxTransmittance, SdTransmittance) %>% unique()

dataLow <- Lightgrid_LowAlt %>% 
  mutate(Zutm_disc = as.integer(Zutm)) %>% 
  group_by(Zutm_disc) %>% 
  mutate(MeanTransmittance = mean(`Period 1`)) %>% 
  mutate(MinTransmittance = min(`Period 1`)) %>% 
  mutate(MaxTransmittance = max(`Period 1`)) %>%
  mutate(SdTransmittance = sd(`Period 1`)) %>% 
  ungroup() %>% 
  select(Zutm_disc, MeanTransmittance, MinTransmittance, MaxTransmittance, SdTransmittance) %>% unique() 


dataLow %>% 
  # filter(Zutm_disc>=0 & Zutm_disc<=60) %>% 
  ggplot(aes(x= Zutm_disc, y= MeanTransmittance)) +
  
  geom_point(data = dataHigh, aes(x = Zutm_disc, y = MeanTransmittance, colour = "High altitude")) +
  geom_line(data = dataHigh, aes(x = Zutm_disc, y = MeanTransmittance, colour = "High altitude")) +
  
  geom_point(data = dataLow, aes(x = Zutm_disc, y = MeanTransmittance, colour = "Low altitude")) +
  geom_line(data = dataLow, aes(x = Zutm_disc, y = MeanTransmittance, colour = "Low altitude")) +
  
  # geom_errorbar(data = dataHigh, aes(ymin=MinTransmittance, ymax=MaxTransmittance), colour = "darkgreen", width=.2) +
  # geom_errorbar(data = dataLow, aes(ymin=MinTransmittance, ymax=MaxTransmittance), colour = "green", width=.2) +
  # geom_errorbar(aes(ymin=MeanTransmittance-SdTransmittance, ymax=MeanTransmittance+SdTransmittance), width=.2)
  
  scale_colour_manual(values = c("High altitude" = "darkgreen",
                                 "Low altitude" = "green")) +
  
  ggtitle("P16 - 2023 - 4ha - Light profile (Grid 10x10m, each 1m in Z)") +
  labs(y="Transmittance", x="Altitude", color = "LiDAR acquisition") +
  xlim(0,70) +
  coord_flip() # coordonnées inversées au départ pour que la ligne suive la hauteur au sol et non le Transmittance
```

# Spatial autoccorelation of the light
```{r}

```


