---
title: "Light maps"
author: "Vincyane Badouard"
date: "`r Sys.Date()`"
output: pdf_document
---

# Packages
```{r, include = F}
library(data.table)
library(tidyverse)
library(sf)
library(terra)
library(corrplot)
library(RColorBrewer)
```

# Spatial objects
```{r}
buffer <- st_as_sf(vect("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/Plot16_4ha_buffer.shp")) # interest zone with buffer
zone <- st_as_sf(vect("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2022/Parcelles_Understory.shp")) # 4 carrés
Understory_stakes <- read.csv("D:/Mes Donnees/PhD/SIG_data/Understory-ALT/Understory_stakes_theoretical_UTM.csv") # stakes on the understory 4ha on P16

```

# Process raw data from AMAPVox
```{r}
# Echo weight by intensity (2 m res):  
path_High <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_buffer_Light_annual_intensity2m.txt"
path_Low <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_LowAlt_4ha_buffer_Light_annual_intensity2m.txt"

path_High_Dry <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_buffer_Light_Dry_intensity2m.txt"
path_High_Wet <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_buffer_Light_Wet_intensity2m.txt"
path_Low_Dry <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_LowAlt_4ha_buffer_Light_Dry_intensity2m.txt"
path_Low_Wet <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_LowAlt_4ha_buffer_Light_Wet_intensity2m.txt"

# UAV
path_UAV2m <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_2m/P16_2023_UAV_4ha_bufferALS_Light_annual_intensity2m.txt"
path_UAV1m <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_1m/P16_2023_UAV_4ha_bufferALS_Light_annual_intensity1m.txt"

# Equal echo weight :
# path_High <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_annual_equalecho2m.txt"
# path_Low <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_buffer_LowAlt_Light_annual_equalecho2m.txt"

# Echo weight by intensity :  
# path_Low_lightgrid <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ.txt"
# path_High_lightgrid <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ.txt"
# path_High <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_annual.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_Dry.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_buffer_Light_Wet.txt"
# path_Low <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_buffer_LowAlt_Light_annual.txt"
# path_2022 <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2022/P16_2022_4ha_buffer_Light_annual.txt"
```

```{r}
ReadLightFile <- function(path, skiplines=10){
  Light <- fread(path, skip=skiplines) # skip the 10 1st lines of header
  Light <- Light[Y != "MEAN"] # enlever les 2 lignes mean
  Light <- janitor::remove_empty(Light,"cols")
  
  Light[, `:=`(X = as.numeric(X) , Y = as.numeric(Y))] # all as numeric
  Light <- Light[!is.na(X)] 
  return(Light)
}

Light_High <- ReadLightFile(path_High)
Light_High
Light_Low <- ReadLightFile(path_Low)

Light_UAV2m <- ReadLightFile(path_UAV2m)
Light_UAV1m <- ReadLightFile(path_UAV1m)

Light_High_Dry <- ReadLightFile(path_High_Dry, skiplines=14)
Light_High_Wet <- ReadLightFile(path_High_Wet, skiplines=14)
Light_Low_Dry <- ReadLightFile(path_Low_Dry, skiplines=14)
Light_Low_Wet <- ReadLightFile(path_Low_Wet, skiplines=14)

```

```{r}
average_Periods <- function(data){
  data <- data[, Transmittance := rowMeans(.SD), .SDcols = c("Period 1", "Period 2", "Period 3", "Period 4", "Period 5", "Period 6")]
}

Light_High_Dry <- average_Periods(Light_High_Dry)
Light_High_Wet <- average_Periods(Light_High_Wet)

Light_Low_Dry <- average_Periods(Light_Low_Dry)
Light_Low_Wet <- average_Periods(Light_Low_Wet)

setnames(Light_High, "Period 1", "Transmittance")
setnames(Light_Low, "Period 1", "Transmittance")
setnames(Light_UAV2m, "Period 1", "Transmittance")
setnames(Light_UAV1m, "Period 1", "Transmittance")
```

## Plot with buffer
```{r}
Plot_Light <- function(Light, acq){
  # Remove edges
  # Light <- Light[X<90 & X> -90 & Y<90 & Y> -90]
  
  lattice::levelplot(Transmittance ~ X*Y, data = Light) # Transmittance
  lattice::levelplot(Z~X*Y, data = Light) # Z = Altitude
  
  ggplot() + 
    geom_raster(data = Light, aes(X, Y, fill = Transmittance)) + 
    scale_fill_gradientn(name = "Light intensity (Transmittance)",
                         colors = terrain.colors(30, rev=T),
                         na.value="white") +
    ggtitle(paste("Paracou P16 4ha -", acq, "- Light map 1m", sep = " ")) +  
    coord_sf()
}

Plot_Light(Light_High, "High altitude")
Plot_Light(Light_Low, "Low altitude")

Plot_Light(Light_UAV2m, "UAV - 2m")
Plot_Light(Light_UAV1m, "UAV - 1m")
```

## Transformation des coordonnées relatives en UTM
On peut retrouver les cooordonnées UTM avec la VOP
```{r}
# Lecture matrice de transformation
VOP <- as.matrix(read.table("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/VOP_P16_9ha.txt"))

Relativ2UTM <- function(data, VOP){
  # Apply inverse of VOP matrix to convert back to UTM
  xyz <- tcrossprod(as.matrix(data[, .(X, Y, Z, c=1)]), solve(VOP))
  data[, `:=`(Xutm = xyz[, 1], Yutm = xyz[, 2], Zutm = xyz[, 3])]
  
  data <- unique(data)
  return(data)
}

Light_HighUTM <- Relativ2UTM(Light_High, VOP)
Light_LowUTM <- Relativ2UTM(Light_Low, VOP)

Light_UAV2mUTM <- Relativ2UTM(Light_UAV2m, VOP)
Light_UAV1mUTM <- Relativ2UTM(Light_UAV1m, VOP)

Light_High_DryUTM <- Relativ2UTM(Light_High_Dry, VOP)
Light_Low_DryUTM <- Relativ2UTM(Light_Low_Dry, VOP)
Light_High_WetUTM <- Relativ2UTM(Light_High_Wet, VOP)
Light_Low_WetUTM <- Relativ2UTM(Light_Low_Wet, VOP)



Plot_LightUTM <- function(Light, zone, buffer){
  ggplot() + 
    geom_point(data = Light, size = 1.28,
               aes(Xutm, Yutm, color = Transmittance)) + 
    scale_color_gradientn(name = "Light intensity (Transmittance)",
                          colors = terrain.colors(30, rev=T),
                          na.value="white") +
    geom_sf(data = sf::st_cast(buffer, "LINESTRING")) + # buffer (to remove after)
    geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
    ggtitle(paste("Paracou P16 4ha - ",deparse(substitute(Light))," - map 1m")) +  
    theme_classic() +
    coord_sf()
}

Plot_LightUTM(Light_HighUTM, zone, buffer)
Plot_LightUTM(Light_LowUTM, zone, buffer)

Plot_LightUTM(Light_UAV2mUTM, zone, buffer)
ggsave(paste("Paracou_P16_4ha_buffer_UAV_Lightmap_1m_intensity_2m.png", sep = ""),
       path = "D:/Mes Donnees/PhD/Figures/lidar/Light_maps/Maps/Intensity_2m",
       width = 18, height = 20, units = "cm", dpi=800, bg="white")

Plot_LightUTM(Light_UAV1mUTM, zone, buffer)
ggsave(paste("Paracou_P16_4ha_buffer_UAV_Lightmap_1m_intensity_1m.png", sep = ""),
       path = "D:/Mes Donnees/PhD/Figures/lidar/Light_maps/Maps/Intensity_2m",
       width = 18, height = 20, units = "cm", dpi=800, bg="white")
```

## Crop on the interest zone
```{r}
Crop <- function(data, zone){
  datasf <- st_as_sf(data, coords = c('Xutm','Yutm'))
  datasf <- st_set_crs(datasf, terra::crs(zone))
  datacrop <- st_intersection(datasf, zone)
  XY <- st_coordinates(datacrop) # stock coordinates
  colnames(XY) <- c("Xutm", "Yutm") 
  st_geometry(datacrop) <- NULL # no more sf object
  datacrop <- cbind(datacrop, XY) # bind XY coord
  return(datacrop)
}

Light_HighUTM <- Crop(Light_HighUTM, zone)
Light_LowUTM <- Crop(Light_LowUTM, zone)

Light_High_DryUTM <- Crop(Light_High_DryUTM, zone)
Light_Low_DryUTM <- Crop(Light_Low_DryUTM, zone)
Light_High_WetUTM <- Crop(Light_High_WetUTM, zone)
Light_Low_WetUTM <- Crop(Light_Low_WetUTM, zone)


Light_UAV2mUTM <- Crop(Light_UAV2mUTM, zone)
Light_UAV1mUTM <- Crop(Light_UAV1mUTM, zone)


Plot_LightUTM <- function(Lightcrop, zone, acq, lim){
  ggplot() + 
    geom_point(data = Lightcrop, size = 1.28,
               aes(Xutm, Yutm, color = Transmittance)) + 
    
    scale_color_gradientn(name = "Light intensity (Transmittance)",
                          colors = terrain.colors(30, rev=T),
                          na.value="white", limits = c(0,lim)) +
    
    geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
    ggtitle(paste("Paracou P16 4ha -", acq, "- Light map 1m", sep = " ")) +  
    theme_classic() +
    coord_sf()
  
  ggsave(paste("Paracou_P16_4ha_", acq, "_Lightmap_1m_intensity_2m.png", sep = ""),
         path = "D:/Mes Donnees/PhD/Figures/lidar/Light_maps/Maps/Intensity_2m",
         width = 18, height = 20, units = "cm", dpi=800, bg="white")
}

Plot_LightUTM(Light_HighUTM, zone, "High altitude", lim=0.26)
Plot_LightUTM(Light_LowUTM, zone, "Low altitude", lim=0.26)

Plot_LightUTM(Light_High_DryUTM, zone, "High altitude - Dry season", lim=0.26)
Plot_LightUTM(Light_Low_DryUTM, zone, "Low altitude - Dry season", lim=0.26)
Plot_LightUTM(Light_High_WetUTM, zone, "High altitude - Wet season", lim=0.26)
Plot_LightUTM(Light_Low_WetUTM, zone, "Low altitude - Wet season", lim=0.26)

Plot_LightUTM(Light_UAV2mUTM, zone, "UAV - 2m", lim= max(Light_UAV2mUTM$Transmittance))
Plot_LightUTM(Light_UAV1mUTM, zone, "UAV - 1m", lim= max(Light_UAV1mUTM$Transmittance))
```

## Write data cropped and in UTM
```{r}
Writedata_txt_csv <- function(data, path){
  fwrite(data, paste(path, ".txt", sep=""), row.names = F, col.names = T)
  write.csv(data, paste(path, ".csv", sep=""))
}

Writedata_txt_csv(Light_HighUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_Light_annual_intensity2m_UTM")
Writedata_txt_csv(Light_LowUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/Intensity_2m/P16_2023_4ha_LowAlt_Light_annual_intensity2m_UTM")

Writedata_txt_csv(Light_High_DryUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_Light_Dry_intensity2m_UTM")
Writedata_txt_csv(Light_Low_DryUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/Intensity_2m/P16_2023_4ha_LowAlt_Light_Dry_intensity2m_UTM")
Writedata_txt_csv(Light_High_WetUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_Light_Wet_intensity2m_UTM")
Writedata_txt_csv(Light_Low_WetUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/Intensity_2m/P16_2023_4ha_LowAlt_Light_Wet_intensity2m_UTM")

Writedata_txt_csv(Light_UAV2mUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_2m/P16_2023_4ha_UAV_Light_annual_intensity2m_UTM")
Writedata_txt_csv(Light_UAV1mUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_1m/P16_2023_4ha_UAV_Light_annual_intensity1m_UTM")

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.txt"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_UTM.txt"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.csv"

# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_UTM.csv"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_UTM.csv"
```

# Load data pré-processed
```{r}
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_annual_1m_UTM.txt"
# "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_HighAlt_4ha_Light_Dry_1m_UTM.txt"

Light_HighUTM <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_Light_annual_intensity2m_UTM.txt")

Light_LowUTM <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/Intensity_2m/P16_2023_4ha_LowAlt_Light_annual_intensity2m_UTM.txt")

Light_UAV2mUTM <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_2m/P16_2023_4ha_UAV_Light_annual_intensity2m_UTM.txt")

Light_UAV1mUTM <- fread("//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_1m/P16_2023_4ha_UAV_Light_annual_intensity1m_UTM.txt")
```

# Maps

# Light differences
leftjoin by coordinates 
difference between `Period.1` of the 2 datasets
```{r}
# names(Light_HighUTM) # "X","Y","Z","Period.1","Zutm","SubPlotCon","Forest","Plot","SubPlot","Projet","Protocole","Xutm","Yutm"

ComputeLightDiff <- function(data1, data2){
  setDT(data1); setDT(data2)
  
  Light <- merge(data1, data2, by=c("Xutm","Yutm","Zutm", "X","Y","Z", "Forest","Plot","SubPlot","Projet","Protocole","SubPlotCon"), all=TRUE, suffixes = c("_data1", "_data2")) # merge datasets
  
  Light[,Lightdiff:= Transmittance_data1-Transmittance_data2] # compute differences (4-3=1, le rslt informe sur la valeur à laquelle on soustrait)
}

Light_HighvsLow <- ComputeLightDiff(Light_HighUTM, Light_LowUTM)
Light_HighvsLow

Light_High_DryvsWet <- ComputeLightDiff(Light_High_DryUTM, Light_High_WetUTM)
Light_Low_DryvsWet <- ComputeLightDiff(Light_Low_DryUTM, Light_Low_WetUTM)

Light_Low_AnnualvsDry <- ComputeLightDiff(Light_LowUTM, Light_Low_DryUTM)
Light_Low_AnnualvsWet <- ComputeLightDiff(Light_LowUTM, Light_Low_WetUTM)


Light_ALShighvsUAV <- ComputeLightDiff(Light_HighUTM, Light_UAV2mUTM)
Light_ALSlowvsUAV <- ComputeLightDiff(Light_LowUTM, Light_UAV2mUTM)
```

## Light differences maps
```{r}
for(L in c("Light_HighvsLow", "Light_High_DryvsWet", "Light_Low_DryvsWet", "Light_Low_AnnualvsDry", "Light_Low_AnnualvsWet", "Light_ALShighvsUAV", "Light_ALSlowvsUAV")){ # 
  
  # hist(L$Lightdiff, breaks = 15)
  
  print(
    ggplot() + 
      geom_point(data = get(L), size = 1.28,
                 aes(Xutm, Yutm, color = Lightdiff)) + 
      scale_colour_gradient2(name = "Light intensity (Transmittance) difference",
                             low="#000066", high="red", mid = "white", midpoint = 0)+
      # scale_colour_steps2(name = "Light intensity (Transmittance) difference",
      #                     low="#000066", high="red", mid = "white", midpoint = 0.00,
      #                     limits = c(Min, Max),
      #                     breaks = c(Min, -0.06, -0.04, -0.02, -0.01, 0.01, 0.02, 0.04, 0.06, Max), # c(Min, -0.08, -0.02, 0.02, 0.06, Max) ; seq(Min, Max, by = 0.02)
      #                     labels = as.character(round(c(Min, -0.06, -0.04, -0.02, -0.01, 0.01, 0.02, 0.04, 0.06, Max), digits = 2))) +
      geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
      ggtitle(paste("Paracou P16 4ha - ", L," 2023 - Light map difference (1m)", sep="")) +  
      theme_classic() +
      theme(legend.title=element_text(size=13), legend.text=element_text(size=13)) +
      coord_sf()
  )
  
  ggsave(paste("Paracou_P16_4ha_", L, "_2023_Lightmap_difference_1m_intensity_2m.png", sep = ""),
         path = "D:/Mes Donnees/PhD/Figures/lidar/Light_maps/Maps/Intensity_2m",
         width = 23, height = 25, units = "cm", dpi=800, bg="white")
}

Min <- min(get(L)$Lightdiff)
Max <- max(get(L)$Lightdiff)

print(
  ggplot() + 
    geom_point(data = get(L), size = 1.28,
               aes(Xutm, Yutm, color = Lightdiff)) + 
    # scale_colour_gradient2(name = "Light intensity (Transmittance) difference", low="#00AFBB", high="#FC4E07", mid = "white", midpoint = 0)+
    scale_colour_steps2(name = "Light intensity (Transmittance) difference",
                        low="#000066", high="red", mid = "white", midpoint = 0.00,
                        limits = c(Min, Max) #,
                        # breaks = seq(-0.1, 0.1, by = 0.025) # c(Min, -0.08, -0.02, 0.02, 0.06, Max) ; seq(Min, Max, by = 0.02)
    ) +
    geom_sf(data = sf::st_cast(zone, "LINESTRING")) + # zone of interest
    ggtitle(paste("Paracou P16 4ha - ", L," 2023 - Light map difference (1m)", sep="")) + 
    theme_classic() +
    coord_sf()
)
# }
```

# Light profiles
```{r}
# Pathgrid_HighAlt <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_equalecho2m.txt"
# Pathgrid_LowAlt <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_equalecho2m.txt"

Pathgrid_HighAlt <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_intensity2m.txt"
Pathgrid_LowAlt <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_intensity2m.txt"

Pathgrid_High_Dry <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_buffer_Light_Dry_Grid10x10m_each1mZ_intensity2m.txt"
Pathgrid_High_Wet <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/Intensity_2m/P16_2023_HighAlt_4ha_buffer_Light_Wet_Grid10x10m_each1mZ_intensity2m.txt"
Pathgrid_Low_Dry <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_LowAlt_4ha_buffer_Light_Dry_Grid10x10m_each1mZ_intensity2m.txt"
Pathgrid_Low_Wet <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/LowAltitudeFlight/Intensity_2m/P16_2023_LowAlt_4ha_buffer_Light_Wet_Grid10x10m_each1mZ_intensity2m.txt"

Pathgrid_UAV2m <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_2m/P16_2023_4ha_UAV_Light_Grid10x10m_each1mZ_intensity2m.txt"
Pathgrid_UAV1m <- "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/HovermapUAV2023/AMAPVox/Intensity_1m/P16_2023_4ha_UAV_Light_Grid10x10m_each1mZ_intensity1m.txt"


Lightgrid_HighAlt <- ReadLightFile(Pathgrid_HighAlt)
Lightgrid_LowAlt <- ReadLightFile(Pathgrid_LowAlt)

Lightgrid_UAV2m <- ReadLightFile(Pathgrid_UAV2m)
Lightgrid_UAV1m <- ReadLightFile(Pathgrid_UAV1m)

Lightgrid_High_Dry <- ReadLightFile(Pathgrid_High_Dry, skiplines=14)
Lightgrid_High_Wet <- ReadLightFile(Pathgrid_High_Wet, skiplines=14)

Lightgrid_Low_Dry <- ReadLightFile(Pathgrid_Low_Dry, skiplines=14)
Lightgrid_Low_Wet <- ReadLightFile(Pathgrid_Low_Wet, skiplines=14)

# Lightgrid_HighUTM <- Relativ2UTM(Lightgrid_HighAlt, VOP)
# Lightgrid_LowUTM <- Relativ2UTM(Lightgrid_LowAlt, VOP)
# 
# Lightgrid_HighUTM <- Crop(Lightgrid_HighUTM, zone)
# Lightgrid_LowUTM <- Crop(Lightgrid_LowUTM, zone)

# Lightgrid_HighUTM ; Lightgrid_LowUTM

# Writedata_txt_csv(Lightgrid_HighUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/HighAltitudeFlight/P16_2023_4ha_HighAlt_Light_Grid10x10m_each1mZ_equalecho2m_UTM")
# Writedata_txt_csv(Lightgrid_LowUTM, "//amap-data.cirad.fr/work/users/VincyaneBadouard/Lidar/ALS2023/lowAltitudeFlight/P16_2023_4ha_LowAlt_Light_Grid10x10m_each1mZ_equalecho2m_UTM")

names(Lightgrid_High_Dry)
```

```{r}
average_Periods <- function(data){
  data <- data[, Transmittance := rowMeans(.SD), .SDcols = c("Period 1", "Period 2", "Period 3", "Period 4", "Period 5", "Period 6")]
}

Lightgrid_High_Dry <- average_Periods(Lightgrid_High_Dry)
Lightgrid_High_Wet <- average_Periods(Lightgrid_High_Wet)

Lightgrid_Low_Dry <- average_Periods(Lightgrid_Low_Dry)
Lightgrid_Low_Wet <- average_Periods(Lightgrid_Low_Wet)

setnames(Lightgrid_HighAlt, "Period 1", "Transmittance")
setnames(Lightgrid_LowAlt, "Period 1", "Transmittance") 
setnames(Lightgrid_UAV2m, "Period 1", "Transmittance")
setnames(Lightgrid_UAV1m, "Period 1", "Transmittance")
```

## Compute Transmitance by altitude
```{r}
names(Lightgrid_HighAlt) # "X","Y","Z",  "Period 1",  "Xutm","Yutm","Zutm"    

compute_Transmitance_byAlt <- function(data){
  data <- setDF(data) %>% 
    mutate(Z_disc = as.integer(Z)) %>% 
    group_by(Z_disc) %>% 
    mutate(MeanTransmittance = mean(Transmittance)) %>% 
    mutate(MinTransmittance = min(Transmittance)) %>% 
    mutate(MaxTransmittance = max(Transmittance)) %>%
    mutate(SdTransmittance = sd(Transmittance)) %>% 
    ungroup() %>% 
    select(Z_disc, MeanTransmittance, MinTransmittance, MaxTransmittance, SdTransmittance) %>% unique()
}

Lightgrid_HighAlt <- compute_Transmitance_byAlt(Lightgrid_HighAlt)
Lightgrid_LowAlt <- compute_Transmitance_byAlt(Lightgrid_LowAlt)

Lightgrid_UAV2m <- compute_Transmitance_byAlt(Lightgrid_UAV2m)
Lightgrid_UAV1m <- compute_Transmitance_byAlt(Lightgrid_UAV1m)

Lightgrid_High_Dry <- compute_Transmitance_byAlt(Lightgrid_High_Dry)
Lightgrid_High_Wet <- compute_Transmitance_byAlt(Lightgrid_High_Wet)

Lightgrid_Low_Dry <- compute_Transmitance_byAlt(Lightgrid_Low_Dry)
Lightgrid_Low_Wet <- compute_Transmitance_byAlt(Lightgrid_Low_Wet)

```

## Light profiles plot
```{r}
plotLightProfiles <- function(data1, data2, data3 = NULL, LegCol, Title){ 
  data1 %>% 
    # filter(Z_disc>=0 & Z_disc<=60) %>% 
    ggplot(aes(x= Z_disc, y= MeanTransmittance)) +
    
    geom_point(data = data1, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[1])) +
    geom_line(data = data1, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[1])) +
    
    geom_point(data = data2, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[2])) +
    geom_line(data = data2, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[2])) +
    
    {if(!is.null(data3))
      list(
        geom_point(data = data3, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[3])),
        geom_line(data = data3, aes(x = Z_disc, y = MeanTransmittance, colour = names(LegCol)[3])) )} +
    
    # geom_errorbar(data = data2, aes(ymin=MinTransmittance, ymax=MaxTransmittance), colour = "darkgreen", width=.2) +
    # geom_errorbar(data = data1, aes(ymin=MinTransmittance, ymax=MaxTransmittance), colour = "green", width=.2) +
    # geom_errorbar(aes(ymin=MeanTransmittance-SdTransmittance, ymax=MeanTransmittance+SdTransmittance), width=.2)
    
    scale_colour_manual(values = LegCol) +
    theme_minimal() +
    ggtitle(paste("P16 - 2023 - 4ha - ", Title, "Light profile (Grid 10x10m, each 1m in Z)")) +
    labs(y="Transmittance", x="Altitude", color = "LiDAR acquisition") +
    xlim(0,70) +
    coord_flip() # coordonnées inversées au départ pour que la ligne suive la hauteur au sol et non le Transmittance
  
  ggsave(paste("ALS2023_P16_4ha_",Title,"_Light_profile_(Grid_10x10m_each_1m_in_Z)_intensity_2m.png"),
         path = "D:/Mes Donnees/PhD/Figures/lidar/Light_maps/Light_profil/Intensity_2m",
         width = 18, height = 20, units = "cm", dpi=800, bg="white")
}

plotLightProfiles(Lightgrid_HighAlt, Lightgrid_LowAlt,
                  LegCol = c("High altitude" = "blue","Low altitude" = "#56B4E9"),
                  Title = "High vs Low")

plotLightProfiles(Lightgrid_HighAlt, Lightgrid_UAV2m,
                  LegCol = c("High altitude" = "blue","UAV" = "#E69F00"),
                  Title = "ALS High vs UAV") 

plotLightProfiles(Lightgrid_LowAlt, Lightgrid_UAV2m,
                  LegCol = c("Low altitude" = "#56B4E9","UAV" = "#E69F00"),
                  Title = "ALS Low vs UAV") 

plotLightProfiles(Lightgrid_HighAlt, Lightgrid_High_Dry, Lightgrid_High_Wet,
                  c("High altitude" = "blue", "Dry season" = "#D55E00","Wet season" = "#0072B2"),
                  "High alt Dry vs Wet")

plotLightProfiles(Lightgrid_LowAlt, Lightgrid_Low_Dry, Lightgrid_Low_Wet,
                  c("Low altitude" = "#56B4E9", "Dry season" = "#D55E00","Wet season" = "#0072B2"),
                  "Low alt Dry vs Wet")
```

# Spatial autoccorelation of the light
```{r}

```


